### 2024-11-14 paper/paper.qmd
Model check

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
data <- read_csv("data/02-analysis_data/analysis_data.csv")

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

#### Load libraries ####
library(tidyverse)
library(rstanarm)

#### Read data ####
data <- read.csv("data/02-analysis_data/analysis_data.csv")

#### Load the saved model ####
first_model <- readRDS("models/Mortality_Model.rds")

#### Create predictions based on model ####
# Get predictions and add them to the dataset
analysis_data <- data %>%
  mutate(predicted_mortality = posterior_predict(first_model) %>% apply(2, mean))

#### Plot Mortality vs Food ####
plot_food <- ggplot(analysis_data, aes(x = Food, y = Mortality)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_line(aes(y = predicted_mortality), color = "red", size = 1) +
  labs(title = "Mortality vs Food", x = "Food", y = "Mortality")

#### Plot Mortality vs Health Expense ####
plot_health_expense <- ggplot(analysis_data, aes(x = Health_expense, y = Mortality)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_line(aes(y = predicted_mortality), color = "red", size = 1) +
  labs(title = "Mortality vs Health Expense", x = "Health Expense", y = "Mortality")

#### Plot Mortality vs Vaccine ####
plot_vaccine <- ggplot(analysis_data, aes(x = Vacinne, y = Mortality)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_line(aes(y = predicted_mortality), color = "red", size = 1) +
  labs(title = "Mortality vs Vaccine", x = "Vaccine", y = "Mortality")

#### Save the plots ####
ggsave("plots/mortality_vs_food.png", plot = plot_food, width = 8, height = 6)
ggsave("plots/mortality_vs_health_expense.png", plot = plot_health_expense, width = 8, height = 6)
ggsave("plots/mortality_vs_vaccine.png", plot = plot_vaccine, width = 8, height = 6)

```
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-14 paper/paper.qmd
update paper

@@
 library(rstanarm)
+library(knitr)
 
 #### Read data ####
-data <- read.csv("data/02-analysis_data/analysis_data.csv")
+data <- read_csv(file = here::here("data/02-analysis_data/analysis_data.csv"))
 
 #### Load the saved model ####
-first_model <- readRDS("models/Mortality_Model.rds")
+first_model <- readRDS(file = here::here("models/Mortality_Model.rds"))
+summary_table <- summary(first_model)
 
-#### Create predictions based on model ####
-# Get predictions and add them to the dataset
-analysis_data <- data %>%
-  mutate(predicted_mortality = posterior_predict(first_model) %>% apply(2, mean))
-
-#### Plot Mortality vs Food ####
-plot_food <- ggplot(analysis_data, aes(x = Food, y = Mortality)) +
-  geom_point(color = "blue", alpha = 0.6) +
-  geom_line(aes(y = predicted_mortality), color = "red", size = 1) +
-  labs(title = "Mortality vs Food", x = "Food", y = "Mortality")
-
-#### Plot Mortality vs Health Expense ####
-plot_health_expense <- ggplot(analysis_data, aes(x = Health_expense, y = Mortality)) +
-  geom_point(color = "blue", alpha = 0.6) +
-  geom_line(aes(y = predicted_mortality), color = "red", size = 1) +
-  labs(title = "Mortality vs Health Expense", x = "Health Expense", y = "Mortality")
-
-#### Plot Mortality vs Vaccine ####
-plot_vaccine <- ggplot(analysis_data, aes(x = Vacinne, y = Mortality)) +
-  geom_point(color = "blue", alpha = 0.6) +
-  geom_line(aes(y = predicted_mortality), color = "red", size = 1) +
-  labs(title = "Mortality vs Vaccine", x = "Vaccine", y = "Mortality")
-
-#### Save the plots ####
-ggsave("plots/mortality_vs_food.png", plot = plot_food, width = 8, height = 6)
-ggsave("plots/mortality_vs_health_expense.png", plot = plot_health_expense, width = 8, height = 6)
-ggsave("plots/mortality_vs_vaccine.png", plot = plot_vaccine, width = 8, height = 6)
+# Display the summary in a formatted table
+kable(summary_table, digits = 2, caption = "Summary of Mortality Model")
 

### 2024-11-16 scripts/03-clean_data.R
script change fit for parquet

@@ -11,6 +11,7 @@
 # Load the necessary library
 library(dplyr)
 library(tidyverse)
+library(arrow)
 # List of file paths
 file_paths <- c("file1.csv", "file2.csv", "file3.csv", "file4.csv")
 
@@ -35,7 +36,7 @@ merged_data <- file1 %>%
   drop_na()
 
 # Save the final merged file
-write.csv(merged_data, "data/02-analysis_data/analysis_data.csv", row.names = FALSE)
+write_parquet(merged_data, "data/02-analysis_data/analysis_data.parquet")
 
 
 

### 2024-11-16 scripts/06-model_data.R

@@ -11,9 +11,10 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(arrow)
 
 #### Read data ####
-analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
 ### Model data ####
 first_model <-

### 2024-11-16 paper/paper.qmd
Model and result sec for the paper

@@
 library(palmerpenguins)
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
@@
 #| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
 ```
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+The goal of our modeling strategy is to investigate how the DDT vaccine coverage, food production index, and current health expenditure per capita (current US$) relate to the under-five mortality rate for each nation.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+We aim to use this model to understand how the above mentioned three factors' impact on child mortality and identify opportunities for improving health outcomes, especially for high mortality rate nations.
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Let $y_i$ be the logrithm of under-five mortality rate for nation $i$. We define the following predictors:
+
+$x_{1i}$: DDT vaccine coverage for nation $i$.
+$x_{2i}$: Food production index for nation $i$.
+$x_{3i}$: Logrithm of current health expenditure per capita (current US$) for nation $i$.
+The model is specified as:
 
 \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
+y_i \mid \mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+u_i &= \alpha + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i} \\
+\alpha &\sim \mbox{Normal}(5, ) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5)\\
+\beta_2 &\sim \mbox{Normal}(0, 2.5)\\
+\beta_3 &\sim \mbox{Normal}(0, 2.5)\\
 \sigma &\sim \mbox{Exponential}(1) 
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We implement this model in R [@citeR] using the rstanarm package [@rstanarm], employing its default priors for predictors and a distribution of \mbox{Normal}(\mu_i, \sigma).
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+We use a multivariate linear model to capture the relationship between the predictors and the response variable.
+This choice is justified by the linear trends observed in the data (see @sec-data).
+The logarithmic transformation of the under-five mortality rate and current health expenditure per capita is applied to stabilize variance and linear relationships, ensuring model validity.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+The Bayesian framework is employed due to its ability to incorporate prior knowledge, improve uncertainty quantification, and handle small sample sizes effectively.
 
@@
 Our results are summarized in @tbl-modelresults.
+
+###Intercept:
+The intercept estimate of 6.643 represents the expected logarithm of the under-five mortality rate when all predictors are held constant at their reference or baseline values (e.g., average food production, average vaccine coverage, and average health expenditure).
+This serves as a baseline for interpreting the effects of the predictors.
+
+###Food:
+Food Production Index:
+The coefficient for the food production index is 0.003, indicating a very slight positive relationship between food production and the under-five mortality rate.
+This naïvely suggests that reducing food production could also reduce mortality, which is counter-intuitive.
+
+However, a closer examination of the confidence interval for this predictor shows that it includes 0.
+This implies that the relationship is not statistically significant, and food production is not strongly related to overall under-five mortality rates.
+This finding suggests that while food production is essential for societal well-being, its immediate impact on reducing child mortality might depend on other factors such as food access, distribution systems, and nutritional quality.
+
+###Vaccine Coverage:
+The coefficient for vaccine coverage is -0.013, meaning that for every 1 percentage point increase in vaccine coverage, the logarithm of the under-five mortality rate decreases by 1.3%, holding other variables constant.
+
+This highlights the critical role of vaccination programs in reducing child mortality.
+For example, increasing vaccine coverage by 10 percentage points could reduce the mortality rate by approximately 13%, emphasizing the importance of robust immunization initiatives.
+
+###Health Expenditure (Log-Transformed):
+The coefficient for the log-transformed health expenditure is -0.529, indicating that a 1% increase in health expenditure per capita is associated with a 0.529 unit decrease in the log of the under-five mortality rate, holding all other predictors constant.
 
@@
 #| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+#| label: tbl-modelresults
 
@@
 library(knitr)
-
-#### Read data ####
-data <- read_csv(file = here::here("data/02-analysis_data/analysis_data.csv"))
+library(arrow)
 
 #### Load the saved model ####
-first_model <- readRDS(file = here::here("models/Mortality_Model.rds"))
-summary_table <- summary(first_model)
+# Load the saved models
+model1 <- readRDS(file = here::here("models/Mortality_Model.rds"))
 
-# Display the summary in a formatted table
-kable(summary_table, digits = 2, caption = "Summary of Mortality Model")
+# Extract a tidy summary of the model
+model_summary <- broom.mixed::tidy(model1, effects = "fixed", conf.int = TRUE) %>%
+  select(term, estimate, std.error, conf.low, conf.high) %>%
+  rename(
+    Term = term,
+    Estimate = estimate,
+    `Std. Error` = std.error,
+    `2.5% CI` = conf.low,
+    `97.5% CI` = conf.high
+  )
+# Display the summary table
+kable(model_summary, digits = 3)
 
 ```
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
 ```
@@
 
-pp_check(first_model) +
+analysis_data <- read_parquet(file = here::here("data/02-analysis_data/analysis_data.parquet"))
+
+pp_check(model1) +
   theme_classic() +
@@
 
-posterior_vs_prior(first_model) +
+posterior_vs_prior(model1) +
   theme_minimal() +
@@
 
-plot(first_model, "trace")
+plot(model1, "trace")
 
-plot(first_model, "rhat")
+plot(model1, "rhat")
 ```

### 2024-11-16 scripts/03-clean_data.R

@@ -33,7 +33,9 @@ merged_data <- file1 %>%
   left_join(file2, by = "country") %>%
   left_join(file3, by = "country") %>%
   left_join(file4, by = "country") %>%
-  drop_na()
+  drop_na() %>%
+  mutate(Health_expense = log(Health_expense))%>%
+  mutate(Mortality = log(Mortality))
 
 # Save the final merged file
 write_parquet(merged_data, "data/02-analysis_data/analysis_data.parquet")

### 2024-11-16 scripts/06-model_data.R

@@ -16,18 +16,16 @@ library(arrow)
 #### Read data ####
 analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = Mortality ~ Food + Health_expense+ Vacinne,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
-
+# Fit the model using stan_glm with custom priors for scaled predictors
+first_model <- stan_glm(
+  formula = Mortality ~ Food + Vacinne + Health_expense,
+  data = analysis_data,
+  family = gaussian(),
+  prior = normal(location = c(0, 0, 0), scale = c(2.5, 2.5, 2.5), autoscale = FALSE),
+  prior_intercept = normal(location = 5, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 853
+)
 
 #### Save model ####
 saveRDS(

### 2024-11-16 paper/paper.qmd
data/response variable

@@
 library(tidyverse)
-library(palmerpenguins)
 
@@
 
-## Overview
+## Overview Check packages
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+For this analysis, we combined four data sets all together into one.
+All these four data sets comes from **Worldbank** open data platform (@worldbank).
+We employed **R** (@citeR), a coding platform to download, clean and conduct statistical analysis.
+Besides, we also utilized R packages **tidyverse** (@tidy), **rstanarm** (@rstanarm), **ggplot2** (@gg), **knitr** (@knitr), **arrow** (@arrow), **here** (@here), and **dpylr** (@dplyr).
+The paper is outlined in github using starter folder provided in **Telling Stories With Data** (@tellingstories).
 
@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+The response variable in our analysis is the mortality rate below 5 year-old children for all countries.
+Here Mortality rate is calculated based on number of deaths of children under 5 years old per 1000 person.
+The histogram of this variable is shown in @fig-mor-1.
+We observe that the data ranges from 0 to 120, with a peak at 10-15.
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+However, this data is extremely right skewed and have several outlyers, thus we decided to perform log transformation to stabilize the variance and in-proving the normality of our dataset for a better regression model latter.
+The distribution after we perform log transformation is shown in @fig-mor-2, here the distribution is approximately normal ranging from 2 to 5 with a peak at 3.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-mor
+#| fig-cap: Histogram for Mortality rate of different countries
+#| fig-subcap: ["Dsitribution before log tranformation", "Distribution after log transformation"]
 #| echo: false
+#| warning: false
+#| message: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
+# Load necessary libraries
+library(ggplot2)
+library(arrow)
+
+# Read the data
+data <- read.csv(file = (here::here('data/01-raw_data/raw_data_Mortality.csv')))
+
+# Create the histogram
+ggplot(data, aes(x = SH.DYN.MORT)) +
+  geom_histogram(binwidth = 10, fill = "skyblue", color = "black", alpha = 0.7) +
+  labs(
+    x = "Under-Five Mortality Rate",
+    y = "Frequency"
+  ) +
   theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+    axis.title = element_text(size = 10)
+  )
+
+analysis_data <- read_parquet(file = (here::here('data/02-analysis_data/analysis_data.parquet')))
+
+# Create the histogram
+ggplot(analysis_data, aes(x = Mortality)) +
+  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black", alpha = 0.7) +
+  labs(
+    x = "Under-Five Mortality Rate",
+    y = "Frequency"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+    axis.title = element_text(size = 10)
+  )
+
 ```
@@
 (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-```
 

### 2024-11-17 paper/paper.qmd
Deleted original words
First draft for data sec

@@
 #| label: fig-mor
-#| fig-cap: Histogram for Mortality rate of different countries
+#| fig-cap: Data Analysis for Mortality Rate of Different Countries
 #| fig-subcap: ["Dsitribution before log tranformation", "Distribution after log transformation"]
@@
 #| message: false
+#| layout-ncol: 2
 
@@
 ggplot(data, aes(x = SH.DYN.MORT)) +
-  geom_histogram(binwidth = 10, fill = "skyblue", color = "black", alpha = 0.7) +
+  geom_histogram(binwidth = 7, fill = "skyblue", color = "black", alpha = 0.7) +
   labs(
@@
     plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-    axis.title = element_text(size = 10)
+    axis.title = element_text(size = 12)
   )
@@
 ggplot(analysis_data, aes(x = Mortality)) +
-  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black", alpha = 0.7) +
+  geom_histogram(binwidth = 0.3, fill = "skyblue", color = "black", alpha = 0.7) +
   labs(
@@
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-Talk way more about it.
-
 ## Predictor variables
 
-Add graphs, tables and text.
+In our mdoel, we have three Preditor variables, namely, Food production index, which is the relative level of agricultural production for each nation compared with the base period 2014-2016 (@worldbank); Current health expenditure per capita (current US$); and DPT vaccine percentage.
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+We will first look at the food production index.
+From @fig-food-1, we see that it ranges from 60 to 180, with a center at 110, and the shape is approximated normal distributed.
+@fig-food-2 is the scatter plot between Food production index and log of Mortality rate, with a best line of fit and standard error.
+We obsereve a slightly positive linear relationship between these two variables.
+Detailed relationship will be studied in @sec-model.
+```{r}
+#| label: fig-food
+#| fig-cap: Data Analysis of Predictor Food Production Index
+#| fig-subcap: ["Histogram of each countries' food production index", "Scatter plot of each countries' Food production index VS. Log of Mortality"]
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
 
-# Model
+# Load necessary libraries
+library(ggplot2)
+library(arrow)
+
+analysis_data <- read_parquet(file = (here::here('data/02-analysis_data/analysis_data.parquet')))
+
+# Create the histogram
+ggplot(analysis_data, aes(x = Food)) +
+  geom_histogram(binwidth = 5, fill = "skyblue", color = "black", alpha = 0.7) +
+  labs(
+    x = "Food Production Index",
+    y = "Frequency"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+    axis.title = element_text(size = 12)
+  )
+
+ggplot(analysis_data, aes(x = Food, y = Mortality)) +
+  geom_point(color = "steelblue", alpha = 0.6) +
+  geom_smooth(method = "lm", color = "darkred", fill = "pink", se = TRUE, alpha = 0.3) +
+  labs(
+    x = "Food Production Index",
+    y = "Mortality Rate",
+  ) +
+  theme_minimal() +
+  theme(
+    axis.title = element_text(size = 12)
+  )
+```
+Next is the Current health expenditure.
+From @fig-health-1, we see that it ranges from 0 to 12000, but the shape of the distribution is extremely skewed to the right.
+Thus we decided to perform a log transformation to stabilize the normal shape.
+@fig-health-2 is the histogram after the log transformation, we see that it now ranges from 0 to 10 with a center at 6 and the shape a approximately normal now.
+@fig-health-3 is a scatter plot between Log of Mortality rate and Log of Current health Expenditure, with a best line of fit and standard error.
+We observe a significant negative linear relationship between these two variables.
+Detailed relationshape will be studied in @sec-model.
+```{r}
+#| label: fig-health
+#| fig-cap: Data Analysis of Predictor Current Health Expenditure
+#| fig-subcap: ["Histogram of each countries' Health Expenditure Per Capita, Measured in US$.
+Before log transformation", "Histogram of each countries' Health Expenditure Per Capita, Measured in US$.
+After log transformation", "Scatter plot of each countries' Health Expenditure Per Capita, Measured in US$ VS.
+Log of Mortality"]
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+
+# Load necessary libraries
+library(ggplot2)
+library(arrow)
+library(knitr)
+
+raw_data <- read.csv(file = (here::here('data/01-raw_data/raw_data_Healthexpense.csv')))
+
+analysis_data <- read_parquet(file = (here::here('data/02-analysis_data/analysis_data.parquet')))
+
+ggplot(raw_data, aes(x = SH.XPD.CHEX.PC.CD)) +
+  geom_histogram(binwidth = 500, fill = "skyblue", color = "black", alpha = 0.7) +
+  labs(
+    x = "Health Expense per capita",
+    y = "Frequency"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+    axis.title = element_text(size = 12)
+  )
+
+# Create the histogram
+ggplot(analysis_data, aes(x = Health_expense)) +
+  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black", alpha = 0.7) +
+  labs(
+    x = "Log of Health Expense per capita",
+    y = "Frequency"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+    axis.title = element_text(size = 12)
+  )
+
+ggplot(analysis_data, aes(x = Health_expense, y = Mortality)) +
+  geom_point(color = "steelblue", alpha = 0.6) +
+  geom_smooth(method = "lm", color = "darkred", fill = "pink", se = TRUE, alpha = 0.3) +
+  labs(
+    x = "Log of Health Expense per capita",
+    y = "Mortality Rate",
+  ) +
+  theme_minimal() +
+  theme(
+    axis.title = element_text(size = 12)
+  )
+```
+Finally is the DPT vaccine percentage.
+From @fig-DPT-1, we see that it ranges from 30 to 100, but the shape of the distribution is skewed to the left, we will discuss this in @sec-results.
+Here we do not have efficient tecnics to stablize the distribution.
+@fig-DPT-2 is a scatter plot between DPT vaccine percentage and Log of Mortality rate, with a best line of fit and standard error.
+We observe a significant negative linear relationship between these two variables.
+Detailed relationship will be studied in @sec-model.
+```{r}
+#| label: fig-DPT
+#| fig-cap: Data Analysis of Predictor DPT Vaccine Percentage
+#| fig-subcap: ["Histogram of each countries' DPT vaccine percentage", "Scatter plot of each countries' DPT Vaccine Percentage VS. Log of Mortality"]
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+
+# Load necessary libraries
+library(ggplot2)
+library(arrow)
+library(knitr)
+library(dplyr)
+
+analysis_data <- read_parquet(file = (here::here('data/02-analysis_data/analysis_data.parquet')))
+
+# Create the histogram
+ggplot(analysis_data, aes(x = Vacinne)) +
+  geom_histogram(binwidth = 5, fill = "skyblue", color = "black", alpha = 0.7) +
+  labs(
+    x = "DPT vaccine percentage",
+    y = "Frequency"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+    axis.title = element_text(size = 12)
+  )
+
+ggplot(analysis_data, aes(x = Vacinne, y = Mortality)) +
+  geom_point(color = "steelblue", alpha = 0.6) +
+  geom_smooth(method = "lm", color = "darkred", fill = "pink", se = TRUE, alpha = 0.3) +
+  labs(
+    x = "DPT vaccine percentage",
+    y = "Mortality Rate",
+  ) +
+  theme_minimal() +
+  theme(
+    axis.title = element_text(size = 12)
+  )
+```
+## Missing Data and Time Inconsistancy
+In our file, the varible Health Expendure is collect for year 2021 while the others is collected in year 2022.
+We done this is because there is n data avliable for year 2022 of this data.
+We believe that these data are recent and so the data collected in year 2021 could still do a good job associating data in 2022.
+Also there are a few data missing for some minor countries. We made the decision to drop them and continue our modeling. We will discuss further implications of these two in @sec-results
+
+# Model {#sec-model}
 
@@
 
-# Results
+# Results {#sec-results}
 

### 2024-11-17 paper/paper.qmd
Intro added

@@
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Junbo Li
+thanks: "Code and data are available at: [https://github.com/Junbo345/Mortality_analysis]."
 date: today
@@
 bibliography: references.bib
+toc: true
 ---
@@
 
-Overview paragraph
+Over the past 30 years, advancements in society and technology have significantly reduced global mortality rates, particularly for children under the age of five.
+The under-five mortality rate declined from 93 deaths per 1,000 live births in 1990 to 37 deaths per 1,000 live births in 2022 (@worldbank).
+This progress represents millions of children who now have better chances of survival, marking a substantial achievement in global health.
 
-Estimand paragraph
+In this study, we examined the under-five mortality rate for each country in 2022, focusing on three key factors: food production index, DPT vaccine coverage, and per-capita health expenditure.
+These factors were chosen because they reflect critical components influencing child mortality.
+Infectious diseases, such as pneumonia, diarrhea, and malaria, remain leading causes of under-five deaths, and these are directly linked to vaccination coverage and healthcare spending.
+Similarly, maternal health during childbirth is influenced by healthcare accessibility and quality, which is often tied to national health expenditures (@who2024childmortality).
+While food production does not directly address infectious disease or maternal health, it reflects a nation’s ability to meet nutritional needs, which is critical for preventing malnutrition—a major underlying cause of child deaths (add supporting explanation and citation).
 
-Results paragraph
+Our findings reveal that increasing vaccine coverage and per-capita health expenditures significantly reduces under-five mortality rates.
+However, food production index does not exhibit a direct impact, suggesting that its role is mediated by other factors such as distribution systems and food quality.
+Despite progress in reducing global mortality, the pace of reduction has slowed considerably since 2015, decreasing from an annual rate of 3.8% during the Millennium Development Goal (MDG) era (2000–2015) to 2.1% during the Sustainable Development Goal (SDG) era (2015–2022) (@unicef2022underfive).
+In 2022 alone, 4.9 million children under the age of five died, equating to approximately 13,400 deaths per day (@unicef2022underfive).
 
-Why it matters paragraph
+Furthermore, the distribution of under-five mortality rates remains highly uneven.
+Developed countries such as Canada and the United States report mortality rates as low as 10 per 1,000 live births, while developing nations—particularly in sub-Saharan Africa—experience rates exceeding 100 per 1,000 live births (@worldbank).
+This disparity highlights the urgent need for targeted interventions in high-mortality regions.
+The factors studied in this paper, such as vaccination and healthcare spending, are actionable and can be effectively improved in these regions through foreign aid and international collaboration.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The remainder of this paper is organized as follows.
+@sec-data describes the data sources, key variables, and preprocessing steps, including transformations applied to address skewness and improve model validity.
+@sec-model outlines the modeling approach, including the choice of predictors, justification of the model structure, and Bayesian implementation.
+@sec-results presents the results of the analysis, highlighting the effects of each predictor on under-five mortality.
+Finally, @sec-diss discusses the implications of the findings, the limitations of the study, and recommendations for future research and policy interventions aimed at reducing child mortality globally.
+
+## Estimand
+The primary estimand of this study is the effect of three predictors—food production index, vaccination coverage (DPT vaccine percentage), and per-capita health expenditure (log-transformed)—on the logarithm of the under-five mortality rate for each country in 2022.
+This analysis seeks to quantify the extent to which these factors contribute to variations in mortality rates across nations.
+By addressing these relationships, the study aims to identify potential intervention points for reducing preventable child deaths and achieving progress toward the Sustainable Development Goals for child survival.
 
@@
 
-# Discussion
+# Discussion {#sec-diss}
 

### 2024-11-17 paper/paper.qmd
style change

@@
 
-###Intercept:
+##Intercept:
 The intercept estimate of 6.643 represents the expected logarithm of the under-five mortality rate when all predictors are held constant at their reference or baseline values (e.g., average food production, average vaccine coverage, and average health expenditure).
@@
 
-###Food:
+##Food:
 Food Production Index:
@@
 
-###Vaccine Coverage:
+##Vaccine Coverage:
 The coefficient for vaccine coverage is -0.013, meaning that for every 1 percentage point increase in vaccine coverage, the logarithm of the under-five mortality rate decreases by 1.3%, holding other variables constant.
@@
 
-###Health Expenditure (Log-Transformed):
+##Health Expenditure (Log-Transformed):
 The coefficient for the log-transformed health expenditure is -0.529, indicating that a 1% increase in health expenditure per capita is associated with a 0.529 unit decrease in the log of the under-five mortality rate, holding all other predictors constant.

### 2024-11-18 paper/paper.qmd
Disscussion

@@
 ```
-```
 
@@
 
-## First discussion point {#sec-first-point}
+## implications
+The UN's Sustainable Development Goal 3.2 aims to reduce the mortality rate of children under five to fewer than 25 per 1,000 live births by 2030 (@who2024).
+While substantial progress has been made, achieving this target within the next six years presents significant challenges.
+This study confirms that while the global trend shows a decrease, the pace of this reduction has slowed since 2015.
+Furthermore, in @sec-data we found significant disparities between countries.
+Some nations already meet the SDG target, but others exhibit mortality rates four times higher than the threshold, predominantly in low-income regions such as sub-Saharan Africa.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+To meet the SDG targets, urgent and focused international collaboration is required.
+Our findings emphasize two critical areas for intervention: vaccine coverage and healthcare spending.
+Vaccination programs have a clear and significant impact on reducing child mortality.
+Therefore, international aid should prioritize making newborn vaccines affordable, ensuring their availability, and facilitating universal access in high-mortality regions.
 
-## Second discussion point
+Healthcare spending presents a more complex challenge.
+It encompasses multiple factors, including a country's economic status, infrastructure development, and access to hospital services.
+Addressing these requires a sustained, long-term commitment.
+Developed countries can play a pivotal role by providing financial support, sharing technological advancements, and fostering knowledge transfer to enhance healthcare systems in developing nations.
+Without addressing these systemic issues, achieving equitable child survival rates will remain elusive.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+## limitation and next steps
+This study has several limitations that should be addressed in future research.
+First, as discussed in Section 2, we lack data from several smaller countries.
+While their populations are minimal, excluding these nations raises potential biases and ethical concerns by not representing their unique challenges.
+Second, the most recent data available is from 2022.
+More current datasets would enable a more accurate analysis, reflecting recent developments in healthcare and child mortality trends.
 
-## Third discussion point
+Additionally, while this study focused on three predictors, the dataset contains other indicators that could potentially enrich the analysis.
+However, many of these indicators suffer from missing or outdated data, limiting their utility.
+Future research should prioritize ensuring data completeness and timeliness to enhance model robustness.
 
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+Lastly, advanced modeling techniques and the incorporation of additional predictors could yield deeper insights.
+Future studies should explore innovative models to capture complex relationships and interactions, thereby providing more nuanced recommendations for policy interventions.
 

### 2024-11-19 paper/paper.qmd
madurement

@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The data used in this study were collected from established international organizations and reflect key metrics in health and socioeconomic development.
+The under-five mortality rate, the primary response variable, is measured as the probability per 1,000 live births that a child will not survive to their fifth birthday, given age-specific mortality rates of the specified year.
+This indicator is developed by the UN Inter-agency Group for Child Mortality Estimation and leverages statistical models to reconcile differences among multiple data sources.
+
+Predictors include:
+
+**Food Production Index**: This metric, provided by the Food and Agriculture Organization, captures the aggregate volume of food production normalized to the 2014-2016 base period.
+It excludes non-nutritive edible items such as coffee and tea and accounts for deductions of intermediate agricultural inputs like seed and feed.
+
+**Health Expenditure per Capita**: Sourced from the World Health Organization’s Global Health Expenditure database, this indicator measures current health spending per capita in current US dollars, including goods and services consumed annually.
+
+**DPT Vaccine Coverage**: This is the percentage of children aged 12-23 months who received three doses of the diphtheria, pertussis, and tetanus vaccine, based on estimates derived from national administrative data and household surveys.
 
@@
 
+## Challenges in Observational Data and Sampling
+The data used in this study, particularly under-five mortality rates, are largely derived from surveys, censuses, and observational data rather than complete vital registration systems.
+Observational data present unique challenges, such as recall bias in surveys, underreporting in regions with weak administrative systems, and inconsistencies in data collection methodologies across countries.
+These issues are particularly acute in low- and middle-income countries (LMICs), where infrastructure limitations impede accurate data capture.
+
+Sampling Strategies
+To mitigate these challenges, the UN Inter-agency Group for Child Mortality Estimation employs a combination of direct and indirect estimation techniques.
+Direct methods rely on household surveys, such as Demographic and Health Surveys (DHS) and Multiple Indicator Cluster Surveys (MICS), which use stratified random sampling to ensure representativeness.
+These surveys ask respondents to recall birth and death histories, which introduces recall bias, particularly for events occurring further in the past.
+
+Indirect methods use model life tables and historical data to extrapolate mortality estimates in the absence of recent or reliable survey data.
+These methods rely on assumptions about fertility, mortality trends, and the demographic structure of the population, which may not hold true in all contexts, particularly in regions experiencing rapid epidemiological transitions or conflict.
+
+## Linkages to Literature
+The challenges and methodologies described align with findings in the literature.
+Alkema and New (2014) emphasized the importance of integrating multiple data sources to improve the reliability of child mortality estimates.
+They highlighted the use of Bayesian hierarchical models to synthesize survey and vital registration data, which is a key component of the UN Inter-agency Group’s methodology.
+
+Moultrie et al.
+(2013) detailed the limitations of indirect estimation methods, particularly their reliance on model assumptions that may not hold in regions undergoing rapid social or economic change.
+Their findings support the use of simulation studies, such as ours, to assess the robustness of mortality estimates under varying data quality scenarios.
+
+## Implications for Policy and Future Research
+The limitations of survey-based and observational data highlight the need for:
+
+Investment in Vital Registration Systems: Increasing the coverage and quality of birth and death registrations in LMICs would reduce reliance on indirect methods and improve the accuracy of mortality estimates.
+Improved Survey Methodologies: Incorporating digital data collection tools and cross-validation with other data sources can reduce recall bias and enhance data reliability.
+Advanced Statistical Techniques: Bayesian hierarchical models and machine learning approaches can integrate diverse data sources and account for biases more effectively than traditional methods.
+Future research should focus on combining survey data with novel data sources, such as satellite imagery and mobile health records, to address data gaps and improve mortality estimation in real time.
+These approaches could revolutionize the monitoring of global health indicators, enabling more targeted and timely interventions.
+(Add refer)
+
 # Model details {#sec-model-details}

### 2024-11-21 paper/paper.qmd
Title, ab, appendix

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Exploring Determinants of Child Mortality: The Role of Vaccination, Healthcare Spending, and Nutrition"
+subtitle: "Findings Reveal Vaccination and Healthcare Spending as Key Drivers of Mortality Reduction"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Child mortality has significantly decreased over the past three decades, but progress remains uneven across regions.
+This study examines the under-five mortality rates of 2022, focusing on the effects of vaccination coverage, health expenditure per capita, and food production indices.
+Using a Bayesian modeling approach, we found that increased vaccination coverage and healthcare spending significantly reduce child mortality, while food production had no direct impact.
+These findings underscore the critical need for targeted interventions in high-mortality regions, emphasizing the importance of vaccination programs and healthcare funding in achieving global child survival goals."
 format: pdf
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+We conducted a posterior predictive check to evaluate how well the model predicts under-five mortality rates.
+@fig-ppcheckandposteriorvsprior-1 illustrates the predictive distribution compared to the observed data.
+The close alignment between observed and predicted values indicates that the model captures the main trends in the data effectively.
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
+This shows that the data significantly updates the prior beliefs for all key parameters.
+The posterior distributions of predictors—vaccine coverage, health expenditure, and food production index—highlight the robustness of their effects in the model.
 
@@
 #| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
+#| layout-ncol: 1
 #| fig-cap: "Examining how the model fits, and is affected by, the data"
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+@fig-stanareyouokay-1 is a trace plot.
+It shows adequate mixing of the Markov Chain Monte Carlo (MCMC) chains, indicating proper convergence.
+Additionally, the Rhat values for all parameters (Figure 6b) are below 1.05, further confirming convergence and the reliability of the parameter estimates.
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-stanareyouokay-2 is a Rhat plot.
+It shows no significant patterns, indicating that the model assumptions hold.
+Additionally, the Bayesian framework effectively quantifies uncertainty, with credible intervals for the predictors providing insight into their respective influences on mortality.
 
@@
 #| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| layout-ncol: 1
+#| fig-width: 8
 

### 2024-11-21 paper/paper.qmd
first feedback from prof

@@
 ---
-title: "Exploring Determinants of Child Mortality: The Role of Vaccination, Healthcare Spending, and Nutrition"
+title: "Exploring Determinants of Child Mortality (2022): The Role of Vaccination, Healthcare Spending, and Nutrition, Measurement robust, figure, no histogram/scatterplot of"
 subtitle: "Findings Reveal Vaccination and Healthcare Spending as Key Drivers of Mortality Reduction"
@@
 Over the past 30 years, advancements in society and technology have significantly reduced global mortality rates, particularly for children under the age of five.
-The under-five mortality rate declined from 93 deaths per 1,000 live births in 1990 to 37 deaths per 1,000 live births in 2022 (@worldbank).
+The under-five mortality rate declined from 93 deaths per 1,000 live births in 1990 to 37 deaths per 1,000 live births in 2022 [@worldbank].
 This progress represents millions of children who now have better chances of survival, marking a substantial achievement in global health.
@@
 Infectious diseases, such as pneumonia, diarrhea, and malaria, remain leading causes of under-five deaths, and these are directly linked to vaccination coverage and healthcare spending.
-Similarly, maternal health during childbirth is influenced by healthcare accessibility and quality, which is often tied to national health expenditures (@who2024childmortality).
+Similarly, maternal health during childbirth is influenced by healthcare accessibility and quality, which is often tied to national health expenditures [@who2024childmortality].
 While food production does not directly address infectious disease or maternal health, it reflects a nation’s ability to meet nutritional needs, which is critical for preventing malnutrition—a major underlying cause of child deaths (add supporting explanation and citation).
@@
 However, food production index does not exhibit a direct impact, suggesting that its role is mediated by other factors such as distribution systems and food quality.
-Despite progress in reducing global mortality, the pace of reduction has slowed considerably since 2015, decreasing from an annual rate of 3.8% during the Millennium Development Goal (MDG) era (2000–2015) to 2.1% during the Sustainable Development Goal (SDG) era (2015–2022) (@unicef2022underfive).
-In 2022 alone, 4.9 million children under the age of five died, equating to approximately 13,400 deaths per day (@unicef2022underfive).
+Despite progress in reducing global mortality, the pace of reduction has slowed considerably since 2015, decreasing from an annual rate of 3.8% during the Millennium Development Goal (MDG) era (2000–2015) to 2.1% during the Sustainable Development Goal (SDG) era (2015–2022) [@unicef2022underfive].
+In 2022 alone, 4.9 million children under the age of five died, equating to approximately 13,400 deaths per day [@unicef2022underfive].
 
@@
 For this analysis, we combined four data sets all together into one.
-All these four data sets comes from **Worldbank** open data platform (@worldbank).
-We employed **R** (@citeR), a coding platform to download, clean and conduct statistical analysis.
-Besides, we also utilized R packages **tidyverse** (@tidy), **rstanarm** (@rstanarm), **ggplot2** (@gg), **knitr** (@knitr), **arrow** (@arrow), **here** (@here), and **dpylr** (@dplyr).
-The paper is outlined in github using starter folder provided in **Telling Stories With Data** (@tellingstories).
+All these four data sets comes from **Worldbank** open data platform [@worldbank].
+We employed **R** [@citeR], a coding platform to download, clean and conduct statistical analysis.
+Besides, we also utilized R packages **tidyverse** [@tidy], **rstanarm** [@rstanarm], **ggplot2** [@gg], **knitr** [@knitr], **arrow** [@arrow], **here** [@here], and **dpylr** [@dplyr].
+The paper is outlined in github using starter folder provided in **Telling Stories With Data** [@tellingstories].
 
@@
 
-In our mdoel, we have three Preditor variables, namely, Food production index, which is the relative level of agricultural production for each nation compared with the base period 2014-2016 (@worldbank); Current health expenditure per capita (current US$); and DPT vaccine percentage.
+In our mdoel, we have three Preditor variables, namely, Food production index, which is the relative level of agricultural production for each nation compared with the base period 2014-2016 [@worldbank]; Current health expenditure per capita (current US$); and DPT vaccine percentage.
 

### 2024-11-23 scripts/00-simulate_data.R
tests

@@ -1,52 +1,39 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates the data that used to predict the mortality rate in 2022,
+# the predictor variables are food production index, health expenses, and DPT vaccine rate
+# Author: Junbo Li
+# Date: 12 NOV 2024
+# Contact: junb.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: None
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
+library(arrow)
 
+# Set seed for reproducibility
+set.seed(123)
 
-#### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+# Generate data
+n <- 100  # Number of observations
+
+# Generate predictors
+HealthExpenses <- runif(n, 0, 4000)  # Health expenses
+DPTVaccineRate <- runif(n, 0, 100)  # DPT vaccine rate
+FoodProductionIndex <- runif(n, 10, 100)  # Food production index
+
+# Create MortalityRate negatively related to the other variables
+MortalityRate <- 100 - 0.01 * HealthExpenses - 0.3 * DPTVaccineRate - 0.2 * FoodProductionIndex + rnorm(n, mean = 0, sd = 5)
+
+# Combine into a data frame
+dataset <- data.frame(
+  MortalityRate = MortalityRate,
+  HealthExpenses = HealthExpenses,
+  DPTVaccineRate = DPTVaccineRate,
+  FoodProductionIndex = FoodProductionIndex
 )
 
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
-
-# Create a dataset by randomly assigning states and parties to divisions
-analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
-  state = sample(
-    states,
-    size = 151,
[…]

### 2024-11-23 scripts/01-test_simulated_data.R

@@ -1,89 +1,54 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests Structure and validity of simulated data to predicted mortality rate
+# Author: Junbo Li
+# Date: 12 NOV 2024
+# Contact: junb.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: None
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
+library(testthat)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+analysis_data <- read_parquet("data/00-simulated_data/simulated_dataset.parquet")
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
-
-
-#### Test data ####
-
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 151 rows.")
-}
-
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
-} else {
-  stop("Test Failed: The dataset does not have 3 columns.")
-}
-
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
-} else {
-  stop("Test Failed: The 'division' column contains duplicate values.")
-}
-
-# Check if the 'state' column contains only valid Australian state names
-valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", 
-                  "Western Australia", "Tasmania", "Northern Territory", 
-                  "Australian Capital Territory")
-
-if (all(analysis_data$state %in% valid_states)) {
-  message("Test Passed: The 'state' column contains only valid Australian state names.")
-} else {
[…]

### 2024-11-23 scripts/04-test_analysis_data.R

@@ -1,69 +1,54 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests Structure and validity of cleaned analysis data to predicted mortality rate
+# Author: Junbo Li
+# Date: 12 NOV 2024
+# Contact: junb.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 library(testthat)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+dataset <- read_parquet(file = here::here("data/02-analysis_data/analysis_data.parquet"))
 
-
-#### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
-
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
-
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
-
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
-})
-
-# Test that the 'state' column is character type
-test_that("'state' is character", {
-  expect_type(analysis_data$state, "character")
-})
-
-# Test that there are no missing values in the dataset
-test_that("no missing values in dataset", {
-  expect_true(all(!is.na(analysis_data)))
-})
-
-# Test that 'division' contains unique values (no duplicates)
-test_that("'division' column contains unique values", {
-  expect_equal(length(unique(analysis_data$division)), 151)
-})
-
-# Test that 'state' contains only valid Australian state or territory names
-valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", "Western Australia", 
-                  "Tasmania", "Northern Territory", "Australian Capital Territory")
-test_that("'state' contains valid Australian state names", {
-  expect_true(all(analysis_data$state %in% valid_states))
-})
-
[…]

### 2024-11-24 paper/paper.qmd
First fine-tune

@@
 ---
-title: "Exploring Determinants of Child Mortality (2022): The Role of Vaccination, Healthcare Spending, and Nutrition, Measurement robust, figure, no histogram/scatterplot of"
+title: "Exploring Determinants of Child Mortality (2022): The Role of Vaccination, Healthcare Spending, and Nutrition figure, no histogram/scatterplot of"
 subtitle: "Findings Reveal Vaccination and Healthcare Spending as Key Drivers of Mortality Reduction"
@@
 
-The data used in this study were collected from established international organizations and reflect key metrics in health and socioeconomic development.
-The under-five mortality rate, the primary response variable, is measured as the probability per 1,000 live births that a child will not survive to their fifth birthday, given age-specific mortality rates of the specified year.
-This indicator is developed by the UN Inter-agency Group for Child Mortality Estimation and leverages statistical models to reconcile differences among multiple data sources.
+The analysis in this study relies on four datasets, each collected and prepared by various UN-affiliated organizations.
+Below, we provide a brief overview of the measurement and transformation processes for these datasets.
+Detailed descriptions and methodological specifics can be found in @sec-methods.
 
-Predictors include:
+### Response variable:
 
-**Food Production Index**: This metric, provided by the Food and Agriculture Organization, captures the aggregate volume of food production normalized to the 2014-2016 base period.
-It excludes non-nutritive edible items such as coffee and tea and accounts for deductions of intermediate agricultural inputs like seed and feed.
+**Mortality Rate**: The child mortality rate is estimated by the UN Inter-agency Group for Child Mortality Estimation (IGME).
+The primary data source is each country's vital registration system.
+These raw data are then processed using a statistical model developed collaboratively by UN agencies and academic institutions to derive robust mortality rate estimates [@worldbank2024M].
 
-**Health Expenditure per Capita**: Sourced from the World Health Organization’s Global Health Expenditure database, this indicator measures current health spending per capita in current US dollars, including goods and services consumed annually.
+### Predictors:
 
-**DPT Vaccine Coverage**: This is the percentage of children aged 12-23 months who received three doses of the diphtheria, pertussis, and tetanus vaccine, based on estimates derived from national administrative data and household surveys.
+**Food Production Index**: This index, prepared by the Food and Agriculture Organization (FAO) of the United Nations, measures aggregate food production relative to the 2014–2016 baseline period.
+It is calculated using price-weighted data on agricultural products for each country on an annual basis [@worldbank_foodindex2024].
+
+**Health Expenditure per Capita**: Data on health expenditure per capita are produced by the World Health Organization (WHO) following the System of Health Accounts 2011 (SHA 2011) framework.
+This framework tracks all health-related spending in a country over a defined time period [@worldbank_health_expenditure_us].
+
+**DPT Vaccine Coverage**: DPT vaccine coverage is jointly estimated by the WHO and UNICEF.
+This metric reflects the percentage of children aged 12–23 months in each country who have received the diphtheria, pertussis (whooping cough), and tetanus (DPT) vaccine.
+Estimates are derived from reports by vaccine service providers and household surveys of vaccination history [@worldbank_dpt_immunization].
 
@@
 
-The response variable in our analysis is the mortality rate below 5 year-old children for all countries.
-Here Mortality rate is calculated based on number of deaths of children under 5 years old per 1000 person.
-The histogram of this variable is shown in @fig-mor-1.
-We observe that the data ranges from 0 to 120, with a peak at 10-15.
+The response variable in this study is the under-five mortality rate, measured as the number of deaths of children under five years old per 1,000 live births.
 
-However, this data is extremely right skewed and have several outlyers, thus we decided to perform log transformation to stabilize the variance and in-proving the normality of our dataset for a better regression model latter.
-The distribution after we perform log transformation is shown in @fig-mor-2, here the distribution is approximately normal ranging from 2 to 5 with a peak at 3.
+The untransformed data range from 0 to 120, with a peak between 10 and 15 deaths per 1,000 live births (@fig-mor-1).
+Due to the presence of extreme values and significant right skewness, a log transformation was applied to stabilize variance and improve normality.
+The transformed data range from 2 to 5, with a peak near 3 (@fig-mor-2).
 
@@
 
-In our mdoel, we have three Preditor variables, namely, Food production index, which is the relative level of agricultural production for each nation compared with the base period 2014-2016 [@worldbank]; Current health expenditure per capita (current US$); and DPT vaccine percentage.
+The three predictors analyzed are:
 
-We will first look at the food production index.
-From @fig-food-1, we see that it ranges from 60 to 180, with a center at 110, and the shape is approximated normal distributed.
-@fig-food-2 is the scatter plot between Food production index and log of Mortality rate, with a best line of fit and standard error.
-We obsereve a slightly positive linear relationship between these two variables.
-Detailed relationship will be studied in @sec-model.
+**Food Production Index:** 
+The Food Production Index measures the relative level of agricultural production for each country compared to the 2014–2016 baseline [@worldbank].
+This index reflects price-weighted data of agricultural products for each country.
+
+Distribution: As shown in @fig-food-1, the index ranges from 60 to 180, with a center around 110 and an approximately normal distribution.
+
+Relationship with Mortality: Scatter plot analysis (@fig-food-2) reveals a slightly positive linear relationship between food production and the log-transformed mortality rate, with a fitted line and confidence intervals.
+This relationship will be explored in greater detail in @sec-model.
 ```{r}
@@
 #| fig-cap: Data Analysis of Predictor Food Production Index
-#| fig-subcap: ["Histogram of each countries' food production index", "Scatter plot of each countries' Food production index VS. Log of Mortality"]
+#| fig-subcap: ["Counts of Food production index for each country, measured as a basis of 2014-1016", "Relationship of each countries' Food production index VS. Log of Mortality rate"]
 #| echo: false
@@
 ```
-Next is the Current health expenditure.
-From @fig-health-1, we see that it ranges from 0 to 12000, but the shape of the distribution is extremely skewed to the right.
-Thus we decided to perform a log transformation to stabilize the normal shape.
-@fig-health-2 is the histogram after the log transformation, we see that it now ranges from 0 to 10 with a center at 6 and the shape a approximately normal now.
-@fig-health-3 is a scatter plot between Log of Mortality rate and Log of Current health Expenditure, with a best line of fit and standard error.
-We observe a significant negative linear relationship between these two variables.
-Detailed relationshape will be studied in @sec-model.
+**Current Health Expenditure:**
+The Current Health Expenditure per Capita (measured in USD) reflects each country's health spending, collected using the System of Health Accounts (SHA 2011) framework.
+
+Distribution: The raw data ranges from $0 to $12,000, as shown in @fig-health-1.
+Due to extreme right skewness, we applied a log transformation to normalize the distribution, resulting in a range of 0 to 10, with a center at 6 (@fig-health-2).
+
+Relationship with Mortality: As illustrated in @fig-health-3, a significant negative linear relationship exists between log-transformed health expenditure and log mortality rate.
+Further analysis is detailed in @sec-model.
 ```{r}
@@
 #| fig-cap: Data Analysis of Predictor Current Health Expenditure
-#| fig-subcap: ["Histogram of each countries' Health Expenditure Per Capita, Measured in US$.
-Before log transformation", "Histogram of each countries' Health Expenditure Per Capita, Measured in US$.
-After log transformation", "Scatter plot of each countries' Health Expenditure Per Capita, Measured in US$ VS.
+#| fig-subcap: ["Counts of each countries' Health Expenditure Per Capita, Measured in US$.
+Before log transformation", "Counts of each countries' Health Expenditure Per Capita, Measured in US$.
+After log transformation", "Relationship of each countries' Health Expenditure Per Capita, Measured in US$ VS.
 Log of Mortality"]
@@
 ```
-Finally is the DPT vaccine percentage.
-From @fig-DPT-1, we see that it ranges from 30 to 100, but the shape of the distribution is skewed to the left, we will discuss this in @sec-results.
-Here we do not have efficient tecnics to stablize the distribution.
-@fig-DPT-2 is a scatter plot between DPT vaccine percentage and Log of Mortality rate, with a best line of fit and standard error.
-We observe a significant negative linear relationship between these two variables.
-Detailed relationship will be studied in @sec-model.
+DPT Vaccine Coverage
+The DPT Vaccine Coverage measures the percentage of children aged 12–23 months who received vaccinations for diphtheria, pertussis, and tetanus.
+
+Distribution: As shown in @fig-DPT-1, coverage ranges from 30% to 100%, with a left-skewed distribution.
+
+Relationship with Mortality: A significant negative association is observed in @fig-DPT-2. This relationship will be analyzed further in @sec-model.
 ```{r}
@@
 #| fig-cap: Data Analysis of Predictor DPT Vaccine Percentage
-#| fig-subcap: ["Histogram of each countries' DPT vaccine percentage", "Scatter plot of each countries' DPT Vaccine Percentage VS. Log of Mortality"]
+#| fig-subcap: ["Counts of each countries' DPT vaccine percentage", "Relationship of each countries' DPT Vaccine Percentage VS. Log of Mortality"]
 #| echo: false
@@
 ## Missing Data and Time Inconsistancy
-In our file, the varible Health Expendure is collect for year 2021 while the others is collected in year 2022.
-We done this is because there is n data avliable for year 2022 of this data.
-We believe that these data are recent and so the data collected in year 2021 could still do a good job associating data in 2022.
-Also there are a few data missing for some minor countries. We made the decision to drop them and continue our modeling. We will discuss further implications of these two in @sec-results
+The health expenditure variable was collected in 2021, while other variables reflect 2022 data.
+This discrepancy is due to unavailability of more recent health expenditure data.
+These 2021 values are assumed to approximate 2022 conditions sufficiently for analysis.
+
+Additionally, some smaller countries lacked complete data and were excluded from the analysis. The implications of these omissions are addressed in @sec-results.
 
@@
 
-##Intercept:
+## Intercept:
 The intercept estimate of 6.643 represents the expected logarithm of the under-five mortality rate when all predictors are held constant at their reference or baseline values (e.g., average food production, average vaccine coverage, and average health expenditure).
@@
 
-##Food:
-Food Production Index:
+## Food Production Index:
 The coefficient for the food production index is 0.003, indicating a very slight positive relationship between food production and the under-five mortality rate.
-This naïvely suggests that reducing food production could also reduce mortality, which is counter-intuitive.
+This naively suggests that reducing food production could also reduce mortality, which is counter-intuitive.
 
@@
 
-##Vaccine Coverage:
+## Vaccine Coverage:
 The coefficient for vaccine coverage is -0.013, meaning that for every 1 percentage point increase in vaccine coverage, the logarithm of the under-five mortality rate decreases by 1.3%, holding other variables constant.
@@
 
-##Health Expenditure (Log-Transformed):
+## Health Expenditure (Log-Transformed):
 The coefficient for the log-transformed health expenditure is -0.529, indicating that a 1% increase in health expenditure per capita is associated with a 0.529 unit decrease in the log of the under-five mortality rate, holding all other predictors constant.
@@
 #| label: tbl-modelresults
+#| tbl-cap: "Outcome predictors of our model"
 
@@
 ## implications
-The UN's Sustainable Development Goal 3.2 aims to reduce the mortality rate of children under five to fewer than 25 per 1,000 live births by 2030 (@who2024).
+The UN's Sustainable Development Goal 3.2 aims to reduce the mortality rate of children under five to fewer than 25 per 1,000 live births by 2030 [@who2024].
 While substantial progress has been made, achieving this target within the next six years presents significant challenges.
@@
 
-# Additional data details
+# Additional data details {#sec-methods}
 

### 2024-11-26 paper/paper.qmd
Model sec second edition

@@
 
-The analysis in this study relies on four datasets, each collected and prepared by various UN-affiliated organizations.
-Below, we provide a brief overview of the measurement and transformation processes for these datasets.
+The analysis in this study relies on four data sets, each collected and prepared by various UN-affiliated organizations.
+Below, we provide a brief overview of the measurement and transformation processes for these data sets.
 Detailed descriptions and methodological specifics can be found in @sec-methods.
@@
 
-The goal of our modeling strategy is to investigate how the DDT vaccine coverage, food production index, and current health expenditure per capita (current US$) relate to the under-five mortality rate for each nation.
-
-We aim to use this model to understand how the above mentioned three factors' impact on child mortality and identify opportunities for improving health outcomes, especially for high mortality rate nations.
-Background details and diagnostics are included in [Appendix -@sec-model-details].
+TOur modeling strategy aims to explore the relationships between DPT vaccine coverage, the food production index, and current health expenditure per capita (in current US$) with the under-five mortality rate across nations.
+By analyzing these factors, we seek to identify opportunities to improve child health outcomes, particularly in countries with high mortality rates.
+Supporting background information and diagnostics are provided in [Appendix -@sec-model-details].
 
@@
 
-Let $y_i$ be the logrithm of under-five mortality rate for nation $i$. We define the following predictors:
+Let $y_i$ be the logarithm of under-five mortality rate for nation $i$. We define the following predictors:
 
@@
 $x_{2i}$: Food production index for nation $i$.
-$x_{3i}$: Logrithm of current health expenditure per capita (current US$) for nation $i$.
+$x_{3i}$: Logarithm of current health expenditure per capita (current US$) for nation $i$.
 The model is specified as:
@@
 u_i &= \alpha + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i} \\
-\alpha &\sim \mbox{Normal}(5, ) \\
+\alpha &\sim \mbox{Normal}(2.5, 1) \\
 \beta_1 &\sim \mbox{Normal}(0, 2.5)\\
@@
 
-We implement this model in R [@citeR] using the rstanarm package [@rstanarm], employing its default priors for predictors and a distribution of \mbox{Normal}(\mu_i, \sigma).
+We implement this model in R [@citeR] using the rstanarm package [@rstanarm].
 
@@
 
-We use a multivariate linear model to capture the relationship between the predictors and the response variable.
-This choice is justified by the linear trends observed in the data (see @sec-data).
-The logarithmic transformation of the under-five mortality rate and current health expenditure per capita is applied to stabilize variance and linear relationships, ensuring model validity.
+A multivariate linear model was chosen to capture the relationships between the predictors (DPT vaccine coverage, food production index, and current health expenditure) and the response variable (under-five mortality rate).
+This choice is supported by observed linear trends in the data (see @sec-data).
+Logarithmic transformations were applied to both the under-five mortality rate and current health expenditure per capita to normalize the distributions and stabilize their linear relationships with the predictors.
 
-The Bayesian framework is employed due to its ability to incorporate prior knowledge, improve uncertainty quantification, and handle small sample sizes effectively.
+Default Bayesian priors were used for the predictors, as they provide a robust starting point.
+For the intercept term, a \mbox{Normal}(2.5, 1) prior was chosen, reflecting the observed distribution of the response variable (see @sec-data).
+
+### Model weakness
+Several limitations are noted in this model:
+
+1. Skewed Predictor Distribution: The distribution of DPT vaccine coverage is skewed left, which may introduce bias. Despite attempts to address this, an ideal solution remains elusive.
+
+2. Potential Collinearity: While each predictor shows an individual linear relationship with the response variable, interdependencies between predictors may affect model accuracy.
+
+3. Simplified Assumptions: The model assumes linearity and does not capture potential non-linear interactions or regional variations that could influence under-five mortality rates.
 

### 2024-11-26 scripts/06-model_data.R

@@ -22,7 +22,7 @@ first_model <- stan_glm(
   data = analysis_data,
   family = gaussian(),
   prior = normal(location = c(0, 0, 0), scale = c(2.5, 2.5, 2.5), autoscale = FALSE),
-  prior_intercept = normal(location = 5, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 2.5, scale = 1, autoscale = TRUE),
   prior_aux = exponential(rate = 1, autoscale = TRUE),
   seed = 853
 )

### 2024-11-26 paper/paper.qmd
Results and discussion second edition

@@
 #| label: fig-mor
-#| fig-cap: Data Analysis for Mortality Rate of Different Countries
-#| fig-subcap: ["Dsitribution before log tranformation", "Distribution after log transformation"]
+#| fig-cap: Data Analysis for Mortality Rate for children under five of Different Countries
+#| fig-subcap: ["Dsitribution of Mortality rate before log tranformation", "Distribution of Mortality rate after log transformation"]
 #| echo: false
@@
 ```
-## Missing Data and Time Inconsistancy
+## Missing Data and Time Inconsistancy {#sec-data-missing}
 The health expenditure variable was collected in 2021, while other variables reflect 2022 data.
@@
 A multivariate linear model was chosen to capture the relationships between the predictors (DPT vaccine coverage, food production index, and current health expenditure) and the response variable (under-five mortality rate).
-This choice is supported by observed linear trends in the data (see @sec-data).
+This choice is supported by observed linear trends between each predictors and response in @sec-data.
 Logarithmic transformations were applied to both the under-five mortality rate and current health expenditure per capita to normalize the distributions and stabilize their linear relationships with the predictors.
@@
 However, a closer examination of the confidence interval for this predictor shows that it includes 0.
-This implies that the relationship is not statistically significant, and food production is not strongly related to overall under-five mortality rates.
-This finding suggests that while food production is essential for societal well-being, its immediate impact on reducing child mortality might depend on other factors such as food access, distribution systems, and nutritional quality.
+This implies that the relationship is not statistically significant, so we conclude that food production is not strongly related to overall under-five mortality rates.
 
@@
 The coefficient for vaccine coverage is -0.013, meaning that for every 1 percentage point increase in vaccine coverage, the logarithm of the under-five mortality rate decreases by 1.3%, holding other variables constant.
-
-This highlights the critical role of vaccination programs in reducing child mortality.
-For example, increasing vaccine coverage by 10 percentage points could reduce the mortality rate by approximately 13%, emphasizing the importance of robust immunization initiatives.
 
@@
 
-## implications
-The UN's Sustainable Development Goal 3.2 aims to reduce the mortality rate of children under five to fewer than 25 per 1,000 live births by 2030 [@who2024].
-While substantial progress has been made, achieving this target within the next six years presents significant challenges.
-This study confirms that while the global trend shows a decrease, the pace of this reduction has slowed since 2015.
-Furthermore, in @sec-data we found significant disparities between countries.
-Some nations already meet the SDG target, but others exhibit mortality rates four times higher than the threshold, predominantly in low-income regions such as sub-Saharan Africa.
+## implications in reducing mortality rate
+The UN's Sustainable Development Goal (SDG) 3.2 aims to reduce the mortality rate of children under five to fewer than 25 per 1,000 live births by 2030 [@who2024].
+While substantial progress has been achieved, reaching this target within the next six years presents significant challenges.
+Although global trends indicate a decline in child mortality rates, the pace of this reduction has slowed since 2015 [@who2024].
+Moreover, as highlighted in @fig-mor-1, significant disparities persist between countries.
+Some nations have already surpassed the SDG target, with mortality rates well below the threshold, whereas others—primarily in low-income regions such as sub-Saharan Africa—experience rates four times higher than the target.
 
-To meet the SDG targets, urgent and focused international collaboration is required.
-Our findings emphasize two critical areas for intervention: vaccine coverage and healthcare spending.
-Vaccination programs have a clear and significant impact on reducing child mortality.
-Therefore, international aid should prioritize making newborn vaccines affordable, ensuring their availability, and facilitating universal access in high-mortality regions.
+To achieve the SDG targets, urgent and focused international collaboration is required. Our findings underscore two critical areas for intervention: vaccine coverage and healthcare spending.
+
+Vaccination programs have a demonstrable and significant impact on reducing child mortality.
+As shown in @fig-DPT-1, the distribution of DPT vaccine coverage is skewed left, indicating that while most countries maintain adequate immunization rates, certain nations lag considerably behind.
+WHO’s analysis of DPT vaccine coverage highlights that low-performing countries are concentrated in Africa, the Eastern Mediterranean, and South-East Asia regions [@who_dtp3_coverage].
+These countries, predominantly developing nations, often face financial constraints that prevent them from funding vaccines for their newborn populations.
+Consequently, international aid is crucial to making DPT vaccines accessible and affordable in these regions.
 
 Healthcare spending presents a more complex challenge.
-It encompasses multiple factors, including a country's economic status, infrastructure development, and access to hospital services.
-Addressing these requires a sustained, long-term commitment.
-Developed countries can play a pivotal role by providing financial support, sharing technological advancements, and fostering knowledge transfer to enhance healthcare systems in developing nations.
-Without addressing these systemic issues, achieving equitable child survival rates will remain elusive.
+It involves multiple factors, including a country’s economic capacity, infrastructure development, and access to healthcare services.
+Addressing these issues requires sustained, long-term efforts.
+Developed nations can play a pivotal role by providing financial support, sharing technological advancements, and facilitating knowledge transfer to strengthen healthcare systems in developing countries.
+Without addressing these systemic challenges, achieving equitable child survival rates will remain an unattainable goal.
 
 ## limitation and next steps
-This study has several limitations that should be addressed in future research.
-First, as discussed in Section 2, we lack data from several smaller countries.
-While their populations are minimal, excluding these nations raises potential biases and ethical concerns by not representing their unique challenges.
-Second, the most recent data available is from 2022.
-More current datasets would enable a more accurate analysis, reflecting recent developments in healthcare and child mortality trends.
+This study has several limitations that should be considered in future research.
+First, as noted in @sec-data-missing, data from several countries is unavailable.
+While these nations have minimal populations, excluding them introduces potential biases and raises ethical concerns by overlooking their unique challenges.
+Second, the most recent data used in this study is from 2022.
+Access to more current data sets would enable a more accurate and relevant analysis, reflecting recent developments in healthcare and child mortality trends.
 
-Additionally, while this study focused on three predictors, the dataset contains other indicators that could potentially enrich the analysis.
-However, many of these indicators suffer from missing or outdated data, limiting their utility.
-Future research should prioritize ensuring data completeness and timeliness to enhance model robustness.
+Additionally, while this study focused on three primary predictors, the dataset includes other indicators that could potentially enrich the analysis.
+Unfortunately, many of these indicators suffer from missing or outdated values, limiting their utility.
+Future research should prioritize data completeness and timeliness to improve model robustness and reliability.
 
-Lastly, advanced modeling techniques and the incorporation of additional predictors could yield deeper insights.
-Future studies should explore innovative models to capture complex relationships and interactions, thereby providing more nuanced recommendations for policy interventions.
+Finally, advanced modeling techniques and the inclusion of additional predictors could yield deeper insights.
+Future studies should explore innovative models to better capture complex relationships and interactions, offering more nuanced recommendations for policy interventions.
 

### 2024-11-27 scripts/03-clean_data.R
script preamble

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the data that is needed for model section
+# Author: Junbo Li
+# Date: 12 NOV 2024
+# Contact: junb.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Raw data in data folder
+# Any other information needed? None
 
 #### Workspace setup ####
 # Load the necessary library

### 2024-11-27 scripts/05-model_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: create and store the Bayesian multi-linear model for this paper
+# Author: Junbo Li
+# Date: 12 NOV 2024
+# Contact: junb.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Cleaned data in data folder
+# Any other information needed? None
 
 
 #### Workspace setup ####

### 2024-11-27 scripts/06-model_evaluation.R
evaluation

@@ -1,0 +1,41 @@
+#### Preamble ####
+# Purpose: Obtain an estimate of test score of our model using LOOCV
+# Author: Junbo Li
+# Date: 12 NOV 2024
+# Contact: junb.li@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Cleaned data and model. 
+# Any other information needed? None
+
+library(arrow)
+
+# Load dataset
+data <- read_parquet(file = here::here('data/02-analysis_data/analysis_data.parquet'))
+dataset <- data[,2:4]  
+target <- data[,5]
+
+n <- nrow(dataset)
+
+predictions <- numeric(n)
+actuals <- target
+
+
+model <- read_rds(file = here::here('models/Mortality_Model.rds'))
+
+for (i in 1:n) {
+  # Split data into training and testing sets
+  train_data <- dataset[-i, ]
+  train_target <- target[-i,]
+  test_data <- dataset[i, , drop = FALSE]
+  
+  # Predict for the left-out observation
+  predictions[i] <- predict(model, newdata = test_data)
+}
+
+# Compute LOOCV R-squared
+ss_total <- sum((actuals - mean(unlist(actuals)))^2)  # Total Sum of Squares
+ss_residual <- sum((actuals - predictions)^2)  # Residual Sum of Squares
+r_squared <- 1 - (ss_residual / ss_total)
+
+# Print the LOOCV R-squared
+cat("LOOCV R-squared:", r_squared, "\n")

### 2024-11-28 paper/paper.qmd
Methodology 2nd edi

@@
 
-# Additional data details {#sec-methods}
+# Sampling Methodology Overview and Evaluation {#sec-methods}
 
-## Challenges in Observational Data and Sampling
-The data used in this study, particularly under-five mortality rates, are largely derived from surveys, censuses, and observational data rather than complete vital registration systems.
-Observational data present unique challenges, such as recall bias in surveys, underreporting in regions with weak administrative systems, and inconsistencies in data collection methodologies across countries.
-These issues are particularly acute in low- and middle-income countries (LMICs), where infrastructure limitations impede accurate data capture.
+## Overview
 
-Sampling Strategies
-To mitigate these challenges, the UN Inter-agency Group for Child Mortality Estimation employs a combination of direct and indirect estimation techniques.
-Direct methods rely on household surveys, such as Demographic and Health Surveys (DHS) and Multiple Indicator Cluster Surveys (MICS), which use stratified random sampling to ensure representativeness.
-These surveys ask respondents to recall birth and death histories, which introduces recall bias, particularly for events occurring further in the past.
+This study explores the relationship between under-five mortality rates and three key predictors: food production index, DPT vaccine coverage, and per capita health expenditure.
+The under-five mortality rate is derived from a combination of vital registration systems, household surveys (e.g., Demographic and Health Surveys and Multiple Indicator Cluster Surveys), and indirect estimation models developed by UN agencies and collaborating academic institutions [@worldbank2024M].
+These data sources aim to address disparities in coverage, particularly in regions where vital registration is incomplete.
 
-Indirect methods use model life tables and historical data to extrapolate mortality estimates in the absence of recent or reliable survey data.
-These methods rely on assumptions about fertility, mortality trends, and the demographic structure of the population, which may not hold true in all contexts, particularly in regions experiencing rapid epidemiological transitions or conflict.
+The food production index, developed by the Food and Agriculture Organization of the United Nations (FAO), quantifies national agricultural output relative to a baseline period (2014–2016) using official national statistics and supplementary surveys [@worldbank_foodindex2024].
+Per capita health expenditure is sourced from the World Health Organization (WHO) using the System of Health Accounts (SHA 2011), which systematically tracks all public and private spending on health services, goods, and infrastructure [@worldbank_health_expenditure_us].
+Lastly, DPT vaccine coverage is estimated through a combination of administrative data from immunization service providers and household surveys, reflecting the proportion of children aged 12–23 months who received three doses of the vaccine [@worldbank_dpt_immunization].
 
-## Linkages to Literature
-The challenges and methodologies described align with findings in the literature.
-Alkema and New (2014) emphasized the importance of integrating multiple data sources to improve the reliability of child mortality estimates.
-They highlighted the use of Bayesian hierarchical models to synthesize survey and vital registration data, which is a key component of the UN Inter-agency Group’s methodology.
+## Population sample frame, and samplng appraoch
 
-Moultrie et al.
-(2013) detailed the limitations of indirect estimation methods, particularly their reliance on model assumptions that may not hold in regions undergoing rapid social or economic change.
-Their findings support the use of simulation studies, such as ours, to assess the robustness of mortality estimates under varying data quality scenarios.
+The target population for the response variable, under-five mortality rate, includes all children under the age of five globally in 2022.
+This metric captures the probability per 1,000 live births that a newborn will die before reaching five years of age, providing a direct measure of child survival across regions and contexts.
+The predictors each represent distinct populations.
+The food production index measures agricultural output for entire national populations, reflecting food availability and potential nutritional impacts.
+Similarly, per capita health expenditure pertains to the entire population within a country, as healthcare spending influences system-wide access and quality.
+DPT vaccine coverage, by contrast, specifically targets the population of children aged 12–23 months, providing insights into immunization rates and public health outreach.
 
-## Implications for Policy and Future Research
-The limitations of survey-based and observational data highlight the need for:
+The sampling frame for each variable aligns with its respective population.
+Under-five mortality rates rely on vital registration systems, which are designed to record all births and deaths.
+In regions where these systems are incomplete, surveys and statistical models supplement the data to ensure national representativeness.
+The food production index uses national agricultural statistics as the primary source, with secondary data from government ministries and international organizations filling gaps [@worldbank_foodindex2024].
+For DPT vaccine coverage, data are collected through administrative immunization records verified by household surveys [@worldbank_dpt_immunization].
+Per capita health expenditure is derived from national accounts and follows the standardized SHA 2011 framework [@worldbank_health_expenditure_us].
 
-Investment in Vital Registration Systems: Increasing the coverage and quality of birth and death registrations in LMICs would reduce reliance on indirect methods and improve the accuracy of mortality estimates.
-Improved Survey Methodologies: Incorporating digital data collection tools and cross-validation with other data sources can reduce recall bias and enhance data reliability.
-Advanced Statistical Techniques: Bayesian hierarchical models and machine learning approaches can integrate diverse data sources and account for biases more effectively than traditional methods.
-Future research should focus on combining survey data with novel data sources, such as satellite imagery and mobile health records, to address data gaps and improve mortality estimation in real time.
-These approaches could revolutionize the monitoring of global health indicators, enabling more targeted and timely interventions.
-(Add refer)
+The sampling approach for these variables combines census-based data collection with stratified sampling in regions lacking comprehensive infrastructure.
+While census methods aim to capture complete population-level data, resource constraints in low- and middle-income countries necessitate reliance on surveys and indirect estimation methods.
+These approaches use stratification to ensure representativeness while accounting for regional disparities.
+
+## Strngth and weakness
+
+The methodological framework demonstrates several strengths.
+First, the primary reliance on census-based or near-census-level data ensures comprehensive coverage for most variables.
+This is particularly evident in high-income regions, where systems such as vital registration achieve over 90% completeness [@WHO2024CRVS].
+Second, the use of supplementary data sources, such as surveys or administrative records, mitigates gaps in the primary datasets, enhancing data quality.
+For instance, the food production index excludes non-nutritive products such as coffee and tea, ensuring a focus on agricultural outputs relevant to nutrition [@worldbank_foodindex2024].
+The use of standardized frameworks, such as SHA 2011 for health expenditure, further supports consistency and comparability across countries.
+
+However, notable weaknesses persist.
+In regions with incomplete infrastructure, such as Sub-Saharan Africa, vital registration systems capture only 44% of deaths, leading to substantial data gaps that require imputation through statistical models [@WHO2024CRVS].
+These models rely on assumptions that may not fully reflect local demographic trends, introducing potential biases.
+Moreover, reliance on household surveys for DPT vaccine coverage and mortality data introduces recall bias, as respondents may inaccurately report past events [@Preston1984AgePattern].
+Even census-based data are susceptible to misreporting, undercounting, or misclassification, which can distort results.
+Furthermore, temporal mismatches among variables (e.g., health expenditure data lagging behind mortality rates) may complicate causal inference and trend analysis.
 

### 2024-11-28 paper/paper.qmd
Sketches

@@
 ---
-title: "Exploring Determinants of Child Mortality (2022): The Role of Vaccination, Healthcare Spending, and Nutrition figure, no histogram/scatterplot of"
+title: "Exploring Determinants of Child Mortality (2022): The Role of Vaccination, Healthcare Spending, and Nutrition"
 subtitle: "Findings Reveal Vaccination and Healthcare Spending as Key Drivers of Mortality Reduction"
@@
 These factors were chosen because they reflect critical components influencing child mortality.
-Infectious diseases, such as pneumonia, diarrhea, and malaria, remain leading causes of under-five deaths, and these are directly linked to vaccination coverage and healthcare spending.
+Infectious diseases, such as pneumonia, diarrhea, and malaria, remain the leading causes of under-five deaths, and these are directly linked to vaccination coverage and healthcare spending.
 Similarly, maternal health during childbirth is influenced by healthcare accessibility and quality, which is often tied to national health expenditures [@who2024childmortality].
@@
 
-Our findings reveal that increasing vaccine coverage and per-capita health expenditures significantly reduces under-five mortality rates.
-However, food production index does not exhibit a direct impact, suggesting that its role is mediated by other factors such as distribution systems and food quality.
+Our findings reveal that increasing vaccine coverage and per-capita health expenditures significantly reduce under-five mortality rates.
+However, the food production index does not exhibit a direct impact, suggesting that its role is mediated by other factors such as distribution systems and food quality.
 Despite progress in reducing global mortality, the pace of reduction has slowed considerably since 2015, decreasing from an annual rate of 3.8% during the Millennium Development Goal (MDG) era (2000–2015) to 2.1% during the Sustainable Development Goal (SDG) era (2015–2022) [@unicef2022underfive].
@@
 
-Furthermore, the distribution of under-five mortality rates remains highly uneven.
-Developed countries such as Canada and the United States report mortality rates as low as 10 per 1,000 live births, while developing nations—particularly in sub-Saharan Africa—experience rates exceeding 100 per 1,000 live births (@worldbank).
+Furthermore, the distribution of mortality rates under five remains highly uneven.
+Developed countries such as Canada and the United States report mortality rates as low as 10 per 1,000 live births while developing nations—particularly in sub-Saharan Africa—experience rates exceeding 100 per 1,000 live births (@worldbank).
 This disparity highlights the urgent need for targeted interventions in high-mortality regions.
@@
 For this analysis, we combined four data sets all together into one.
-All these four data sets comes from **Worldbank** open data platform [@worldbank].
-We employed **R** [@citeR], a coding platform to download, clean and conduct statistical analysis.
+All these four datasets come from **Worldbank** open data platform [@worldbank].
+We employed **R** [@citeR], a coding platform to download, clean, and conduct statistical analysis.
 Besides, we also utilized R packages **tidyverse** [@tidy], **rstanarm** [@rstanarm], **ggplot2** [@gg], **knitr** [@knitr], **arrow** [@arrow], **here** [@here], and **dpylr** [@dplyr].
-The paper is outlined in github using starter folder provided in **Telling Stories With Data** [@tellingstories].
+The paper is outlined in Git Hub using the starter folder provided in **Telling Stories With Data** [@tellingstories].
 
@@
 **Food Production Index**: This index, prepared by the Food and Agriculture Organization (FAO) of the United Nations, measures aggregate food production relative to the 2014–2016 baseline period.
-It is calculated using price-weighted data on agricultural products for each country on an annual basis [@worldbank_foodindex2024].
+It is calculated using price-weighted data on agricultural products for each country yearly.
+[@worldbank_foodindex2024].
 
@@
 **DPT Vaccine Coverage**: DPT vaccine coverage is jointly estimated by the WHO and UNICEF.
-This metric reflects the percentage of children aged 12–23 months in each country who have received the diphtheria, pertussis (whooping cough), and tetanus (DPT) vaccine.
+This metric reflects the percentage of children aged 12–23 months in each country who have received diphtheria, pertussis (whooping cough), and tetanus (DPT) vaccine.
 Estimates are derived from reports by vaccine service providers and household surveys of vaccination history [@worldbank_dpt_immunization].
@@
 
-The untransformed data range from 0 to 120, with a peak between 10 and 15 deaths per 1,000 live births (@fig-mor-1).
-Due to the presence of extreme values and significant right skewness, a log transformation was applied to stabilize variance and improve normality.
+The Original data range from 0 to 120, with a peak between 10 and 15 deaths per 1,000 live births (@fig-mor-1).
+Due to extreme values and significant right skewness, a log transformation was applied to stabilize variance and improve normality.
 The transformed data range from 2 to 5, with a peak near 3 (@fig-mor-2).
@@
 **Food Production Index:** 
-The Food Production Index measures the relative level of agricultural production for each country compared to the 2014–2016 baseline [@worldbank].
-This index reflects price-weighted data of agricultural products for each country.
-
-Distribution: As shown in @fig-food-1, the index ranges from 60 to 180, with a center around 110 and an approximately normal distribution.
-
-Relationship with Mortality: Scatter plot analysis (@fig-food-2) reveals a slightly positive linear relationship between food production and the log-transformed mortality rate, with a fitted line and confidence intervals.
-This relationship will be explored in greater detail in @sec-model.
+The Food Production Index measures the relative level of agricultural production for each country, using the 2014–2016 baseline as a reference point [@worldbank].
+This index is price-weighted and reflects agricultural productivity across countries.
+As illustrated in @fig-food-1, the index ranges from 60 to 180, with a mean of around 110 and an approximately normal distribution.
+Scatter plot analysis in @fig-food-2 reveals a slightly positive linear relationship between the Food Production Index and the log-transformed mortality rate.
+This suggests that as agricultural production increases, mortality rates tend to rise modestly, a pattern that warrants further exploration in @sec-model.
 ```{r}
@@
 **Current Health Expenditure:**
-The Current Health Expenditure per Capita (measured in USD) reflects each country's health spending, collected using the System of Health Accounts (SHA 2011) framework.
-
-Distribution: The raw data ranges from $0 to $12,000, as shown in @fig-health-1.
-Due to extreme right skewness, we applied a log transformation to normalize the distribution, resulting in a range of 0 to 10, with a center at 6 (@fig-health-2).
-
-Relationship with Mortality: As illustrated in @fig-health-3, a significant negative linear relationship exists between log-transformed health expenditure and log mortality rate.
-Further analysis is detailed in @sec-model.
+The Current Health Expenditure per Capita captures health spending in US dollars, calculated under the System of Health Accounts (SHA 2011) framework.
+The raw data range from $0 to $12,000 and exhibit extreme right skewness, as shown in @fig-health-1.
+To address this, a log transformation was applied, resulting in a normalized range from 0 to 10, with a central tendency near 6 (@fig-health-2).
+Analysis of the relationship between log-transformed health expenditure and log mortality rate (@fig-health-3) demonstrates a significant negative linear trend, indicating that higher health expenditures are associated with lower mortality rates.
+This relationship is examined in greater detail in @sec-model.
 ```{r}
@@
 ```
-DPT Vaccine Coverage
-The DPT Vaccine Coverage measures the percentage of children aged 12–23 months who received vaccinations for diphtheria, pertussis, and tetanus.
-
-Distribution: As shown in @fig-DPT-1, coverage ranges from 30% to 100%, with a left-skewed distribution.
-
-Relationship with Mortality: A significant negative association is observed in @fig-DPT-2. This relationship will be analyzed further in @sec-model.
+**DPT Vaccine Coverage:**
+DPT vaccine coverage rate reflects the percentage of children aged 12–23 months who received vaccinations against diphtheria, pertussis, and tetanus.
+Coverage ranges from 30% to 100%, with a left-skewed distribution, as depicted in @fig-DPT-1.
+Analysis of the association between vaccine coverage and mortality rates (@fig-DPT-2) shows a strong negative relationship, suggesting that increased vaccine coverage is correlated with lower mortality rates.
+This finding will be explored further in @sec-model.
 ```{r}
@@
 The health expenditure variable was collected in 2021, while other variables reflect 2022 data.
-This discrepancy is due to unavailability of more recent health expenditure data.
+This discrepancy is due to the unavailability of more recent health expenditure data.
 These 2021 values are assumed to approximate 2022 conditions sufficiently for analysis.
@@
 
-TOur modeling strategy aims to explore the relationships between DPT vaccine coverage, the food production index, and current health expenditure per capita (in current US$) with the under-five mortality rate across nations.
+Our modeling strategy aims to explore the relationships between DPT vaccine coverage, the food production index, and current health expenditure per capita (in current US$) with the under-five mortality rate across nations.
 By analyzing these factors, we seek to identify opportunities to improve child health outcomes, particularly in countries with high mortality rates.
@@
 
-A multivariate linear model was chosen to capture the relationships between the predictors (DPT vaccine coverage, food production index, and current health expenditure) and the response variable (under-five mortality rate).
-This choice is supported by observed linear trends between each predictors and response in @sec-data.
-Logarithmic transformations were applied to both the under-five mortality rate and current health expenditure per capita to normalize the distributions and stabilize their linear relationships with the predictors.
+A Bayesian multivariate linear model was chosen to capture the relationships between the predictors (DPT vaccine coverage, food production index, and current health expenditure) and the response variable (under-five mortality rate).
+This choice is supported by observed linear trends between each predictor and response in @sec-data.
+Logarithmic transformations were applied to the under-five mortality rate and current health expenditure per capita to normalize the distributions and stabilize their linear relationships with the predictors.
 
@@
 The intercept estimate of 6.643 represents the expected logarithm of the under-five mortality rate when all predictors are held constant at their reference or baseline values (e.g., average food production, average vaccine coverage, and average health expenditure).
-This serves as a baseline for interpreting the effects of the predictors.
 
@@
 
-## implications in reducing mortality rate
+## Implications in reducing mortality rate
 The UN's Sustainable Development Goal (SDG) 3.2 aims to reduce the mortality rate of children under five to fewer than 25 per 1,000 live births by 2030 [@who2024].
@@
 
-## limitation and next steps
+## Limitation and next steps
 This study has several limitations that should be considered in future research.
@@
 
-## Population sample frame, and samplng appraoch
+## Population sample frame, and sampling approach
 
@@
 
-## Strngth and weakness
+## Strength and weakness
 

### 2024-11-28 paper/paper.qmd
Second edition

@@
 ---
-title: "Exploring Determinants of Child Mortality (2022): The Role of Vaccination, Healthcare Spending, and Nutrition"
+title: "Exploring Determinants of Child Mortality Globally (2022): The Role of Vaccination, Healthcare Spending, and Nutrition"
 subtitle: "Findings Reveal Vaccination and Healthcare Spending as Key Drivers of Mortality Reduction"
@@
 Over the past 30 years, advancements in society and technology have significantly reduced global mortality rates, particularly for children under the age of five.
-The under-five mortality rate declined from 93 deaths per 1,000 live births in 1990 to 37 deaths per 1,000 live births in 2022 [@worldbank].
+The under-five mortality rate declined from 93 deaths per 1,000 live births in 1990 to 37 deaths per 1,000 live births in 2022 [@worldbank2024M].
 This progress represents millions of children who now have better chances of survival, marking a substantial achievement in global health.
@@
 For this analysis, we combined four data sets all together into one.
-All these four datasets come from **Worldbank** open data platform [@worldbank].
+All these four datasets are downloaded through **Worldbank Open Data** API [@worldbank].
 We employed **R** [@citeR], a coding platform to download, clean, and conduct statistical analysis.
-Besides, we also utilized R packages **tidyverse** [@tidy], **rstanarm** [@rstanarm], **ggplot2** [@gg], **knitr** [@knitr], **arrow** [@arrow], **here** [@here], and **dpylr** [@dplyr].
+Besides, we also utilized R packages **tidyverse** [@tidy], **rstanarm** [@rstanarm], **ggplot2** [@gg], **knitr** [@knitr], **arrow** [@arrow], **here** [@here], **modelsummary** [@ms], and **dpylr** [@dplyr].
 The paper is outlined in Git Hub using the starter folder provided in **Telling Stories With Data** [@tellingstories].
@@
 The primary data source is each country's vital registration system.
-These raw data are then processed using a statistical model developed collaboratively by UN agencies and academic institutions to derive robust mortality rate estimates [@worldbank2024M].
+In countries such system is lacking, household surveys are conducted and then processed using a statistical model developed collaboratively by UN agencies and academic institutions to derive robust mortality rate estimates [@worldbank2024M].
 
@@
 #| label: fig-mor
-#| fig-cap: Data Analysis for Mortality Rate for children under five of Different Countries
+#| fig-cap: Data Analysis for Mortality Rate for children under five of different countries.
+The histogram on the top left shows that the original distribution is skewed to the right.
+The histogram on the top right is the distribution after log transformation and is approximately normal.
 #| fig-subcap: ["Dsitribution of Mortality rate before log tranformation", "Distribution of Mortality rate after log transformation"]
@@
 #| label: fig-food
-#| fig-cap: Data Analysis of Predictor Food Production Index
+#| fig-cap: Data Analysis of Predictor Food Production Index.
+The histogram on the left shows a approximately normal distrubution.
+The scatterplot on the right indicates there is a slight positive relationship between Food Production Index and Mortality rate
 #| fig-subcap: ["Counts of Food production index for each country, measured as a basis of 2014-1016", "Relationship of each countries' Food production index VS. Log of Mortality rate"]
@@
 #| label: fig-health
-#| fig-cap: Data Analysis of Predictor Current Health Expenditure
+#| fig-cap: Data Analysis of health expenses per capita.
+The histogram on the top left shows that the original distribution is skewed to the right.
+The histogram on the top right is the distribution after log transformation and is approximately normal.The scatterplot on the bottom left indicates there is a negative relationship between log of health expenses and Mortality rate.
 #| fig-subcap: ["Counts of each countries' Health Expenditure Per Capita, Measured in US$.
@@
 #| label: fig-DPT
-#| fig-cap: Data Analysis of Predictor DPT Vaccine Percentage
+#| fig-cap: Data Analysis of DPT vaccine coverage.
+The histogram on the left shows that the distribution DPT vaccine coverage of is skewed to the left.
+The scatterplot on the right indicates there is a negative relationship between DPT vaccine coverage and Mortality rate.
 #| fig-subcap: ["Counts of each countries' DPT vaccine percentage", "Relationship of each countries' DPT Vaccine Percentage VS. Log of Mortality"]
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Outcome predictors of our model"
+#| tbl-cap: "Explanatory model of Mortality rate for each country based on food production index, DPT vaccine coverage, and health expense"
 
@@
 library(arrow)
+library(modelsummary)
 
@@
 
-# Extract a tidy summary of the model
-model_summary <- broom.mixed::tidy(model1, effects = "fixed", conf.int = TRUE) %>%
-  select(term, estimate, std.error, conf.low, conf.high) %>%
-  rename(
-    Term = term,
-    Estimate = estimate,
-    `Std. Error` = std.error,
-    `2.5% CI` = conf.low,
-    `97.5% CI` = conf.high
-  )
-# Display the summary table
-kable(model_summary, digits = 3)
+modelsummary(list("Mortality rate model" = model1), fmt = 4, statistic = c("({conf.low}, {conf.high})"))
 

### 2024-11-28 paper/paper.qmd
styler change

@@
 library(tidyverse)
-
 ```
@@
 # Read the data
-data <- read.csv(file = (here::here('data/01-raw_data/raw_data_Mortality.csv')))
+data <- read.csv(file = (here::here("data/01-raw_data/raw_data_Mortality.csv")))
 
@@
 
-analysis_data <- read_parquet(file = (here::here('data/02-analysis_data/analysis_data.parquet')))
+analysis_data <- read_parquet(file = (here::here("data/02-analysis_data/analysis_data.parquet")))
 
@@
   )
-
 ```
@@
 
-raw_data <- read.csv(file = (here::here('data/01-raw_data/raw_data_Healthexpense.csv')))
+raw_data <- read.csv(file = (here::here("data/01-raw_data/raw_data_Healthexpense.csv")))
 
-analysis_data <- read_parquet(file = (here::here('data/02-analysis_data/analysis_data.parquet')))
+analysis_data <- read_parquet(file = (here::here("data/02-analysis_data/analysis_data.parquet")))
 
@@
 modelsummary(list("Mortality rate model" = model1), fmt = 4, statistic = c("({conf.low}, {conf.high})"))
-
 ```

### 2024-11-28 scripts/00-simulate_data.R

@@ -17,12 +17,12 @@ library(arrow)
 set.seed(123)
 
 # Generate data
-n <- 100  # Number of observations
+n <- 100 # Number of observations
 
 # Generate predictors
-HealthExpenses <- runif(n, 0, 4000)  # Health expenses
-DPTVaccineRate <- runif(n, 0, 100)  # DPT vaccine rate
-FoodProductionIndex <- runif(n, 10, 100)  # Food production index
+HealthExpenses <- runif(n, 0, 4000) # Health expenses
+DPTVaccineRate <- runif(n, 0, 100) # DPT vaccine rate
+FoodProductionIndex <- runif(n, 10, 100) # Food production index
 
 # Create MortalityRate negatively related to the other variables
 MortalityRate <- 100 - 0.01 * HealthExpenses - 0.3 * DPTVaccineRate - 0.2 * FoodProductionIndex + rnorm(n, mean = 0, sd = 5)

### 2024-11-28 scripts/01-test_simulated_data.R

@@ -17,38 +17,44 @@ analysis_data <- read_parquet("data/00-simulated_data/simulated_dataset.parquet"
 
 #### Test cases ####
 test_that("Dataset structure and value checks", {
-  
   # Test if the dataset has 100 rows
   expect_equal(nrow(dataset), 100, info = "The dataset should have 100 rows.")
-  
+
   # Test if the dataset has 4 columns
   expect_equal(ncol(dataset), 4, info = "The dataset should have 4 columns.")
-  
+
   # Test if 'MortalityRate' values are within the expected range
-  expect_true(all(dataset$MortalityRate >= 5 & dataset$MortalityRate <= 100), 
-              info = "'MortalityRate' values should be within the range 5 to 100.")
-  
+  expect_true(all(dataset$MortalityRate >= 5 & dataset$MortalityRate <= 100),
+    info = "'MortalityRate' values should be within the range 5 to 100."
+  )
+
   # Test if 'HealthExpenses' values are within the expected range
-  expect_true(all(dataset$HealthExpenses >= 0 & dataset$HealthExpenses <= 4000), 
-              info = "'HealthExpenses' values should be within the range 0 to 4000.")
-  
+  expect_true(all(dataset$HealthExpenses >= 0 & dataset$HealthExpenses <= 4000),
+    info = "'HealthExpenses' values should be within the range 0 to 4000."
+  )
+
   # Test if 'DPTVaccineRate' values are within the expected range
-  expect_true(all(dataset$DPTVaccineRate >= 0 & dataset$DPTVaccineRate <= 100), 
-              info = "'DPTVaccineRate' values should be within the range 0 to 100.")
-  
+  expect_true(all(dataset$DPTVaccineRate >= 0 & dataset$DPTVaccineRate <= 100),
+    info = "'DPTVaccineRate' values should be within the range 0 to 100."
+  )
+
   # Test if 'FoodProductionIndex' values are within the expected range
-  expect_true(all(dataset$FoodProductionIndex >= 10 & dataset$FoodProductionIndex <= 100), 
-              info = "'FoodProductionIndex' values should be within the range 10 to 100.")
-  
+  expect_true(all(dataset$FoodProductionIndex >= 10 & dataset$FoodProductionIndex <= 100),
+    info = "'FoodProductionIndex' values should be within the range 10 to 100."
+  )
+
   # Test if there are no missing values
   expect_true(all(!is.na(dataset)), info = "The dataset should not contain any missing values.")
-  
[…]

### 2024-11-28 scripts/02-download_data.R

@@ -18,9 +18,7 @@ data_DPT <- WDI(indicator = "SH.IMM.IDPT", start = 2022, end = 2022)
 data_Healthexpense <- WDI(indicator = "SH.XPD.CHEX.PC.CD", start = 2021, end = 2021)
 
 #### Save data ####
-write_csv(data_Food, "data/01-raw_data/raw_data_Food.csv") 
+write_csv(data_Food, "data/01-raw_data/raw_data_Food.csv")
 write_csv(data_Mortality, "data/01-raw_data/raw_data_Mortality.csv")
 write_csv(data_DPT, "data/01-raw_data/raw_data_DPT.csv")
 write_csv(data_Healthexpense, "data/01-raw_data/raw_data_Healthexpense.csv")
-
-         

### 2024-11-28 scripts/03-clean_data.R

@@ -16,16 +16,16 @@ library(arrow)
 file_paths <- c("file1.csv", "file2.csv", "file3.csv", "file4.csv")
 
 # Read and select columns from each file, renaming the unique column
-file1 <- read.csv('data/01-raw_data/raw_data_Food.csv') %>%
+file1 <- read.csv("data/01-raw_data/raw_data_Food.csv") %>%
   select(country, Food = AG.PRD.FOOD.XD)
 
-file4 <- read.csv('data/01-raw_data/raw_data_Mortality.csv') %>%
+file4 <- read.csv("data/01-raw_data/raw_data_Mortality.csv") %>%
   select(country, Mortality = SH.DYN.MORT)
 
-file3 <- read.csv('data/01-raw_data/raw_data_DPT.csv') %>%
+file3 <- read.csv("data/01-raw_data/raw_data_DPT.csv") %>%
   select(country, Vacinne = SH.IMM.IDPT)
 
-file2 <- read.csv('data/01-raw_data/raw_data_Healthexpense.csv') %>%
+file2 <- read.csv("data/01-raw_data/raw_data_Healthexpense.csv") %>%
   select(country, Health_expense = SH.XPD.CHEX.PC.CD)
 
 # Merge all data frames by 'common_column'
@@ -34,12 +34,8 @@ merged_data <- file1 %>%
   left_join(file3, by = "country") %>%
   left_join(file4, by = "country") %>%
   drop_na() %>%
-  mutate(Health_expense = log(Health_expense))%>%
+  mutate(Health_expense = log(Health_expense)) %>%
   mutate(Mortality = log(Mortality))
 
 # Save the final merged file
 write_parquet(merged_data, "data/02-analysis_data/analysis_data.parquet")
-
-
-
-

### 2024-11-28 scripts/04-test_analysis_data.R

@@ -17,38 +17,44 @@ dataset <- read_parquet(file = here::here("data/02-analysis_data/analysis_data.p
 
 #### Test cases ####
 test_that("Dataset structure and value checks", {
-  
   # Test if the dataset has 100 rows
   expect_equal(nrow(dataset), 182, info = "The dataset should have 100 rows.")
-  
+
   # Test if the dataset has 4 columns
   expect_equal(ncol(dataset), 5, info = "The dataset should have 4 columns.")
-  
+
   # Test if 'MortalityRate' values are within the expected range
-  expect_true(all(dataset$Mortality >= 0 & dataset$Mortality <= 100), 
-              info = "'MortalityRate' values should be within the range 0 to 100.")
-  
+  expect_true(all(dataset$Mortality >= 0 & dataset$Mortality <= 100),
+    info = "'MortalityRate' values should be within the range 0 to 100."
+  )
+
   # Test if 'HealthExpenses' values are within the expected range
-  expect_true(all(dataset$Health_expense >= 0), 
-              info = "'HealthExpenses' values should be larger than 0.")
-  
+  expect_true(all(dataset$Health_expense >= 0),
+    info = "'HealthExpenses' values should be larger than 0."
+  )
+
   # Test if 'DPTVaccineRate' values are within the expected range
-  expect_true(all(dataset$Vacinne >= 0 & dataset$Vacinne <= 100), 
-              info = "'DPTVaccineRate' values should be within the range 0 to 100.")
-  
+  expect_true(all(dataset$Vacinne >= 0 & dataset$Vacinne <= 100),
+    info = "'DPTVaccineRate' values should be within the range 0 to 100."
+  )
+
   # Test if 'FoodProductionIndex' values are within the expected range
-  expect_true(all(dataset$Food >= 0 & dataset$Food <= 500), 
-              info = "'FoodProductionIndex' values should be within the range 0 to 500.")
-  
+  expect_true(all(dataset$Food >= 0 & dataset$Food <= 500),
+    info = "'FoodProductionIndex' values should be within the range 0 to 500."
+  )
+
   # Test if there are no missing values
   expect_true(all(!is.na(dataset)), info = "The dataset should not contain any missing values.")
-  
+
   # Test if 'MortalityRate' has at least two unique values
-  expect_true(length(unique(dataset$Mortality)) >= 2, 
-              info = "'MortalityRate' should have at least two unique values.")
-  
[…]

### 2024-11-28 scripts/05-model_data.R

@@ -32,5 +32,3 @@ saveRDS(
   first_model,
   file = "models/Mortality_Model.rds"
 )
-
-

### 2024-11-28 scripts/06-model_evaluation.R

@@ -4,15 +4,15 @@
 # Date: 12 NOV 2024
 # Contact: junb.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Cleaned data and model. 
+# Pre-requisites: Cleaned data and model.
 # Any other information needed? None
 
 library(arrow)
 
 # Load dataset
-data <- read_parquet(file = here::here('data/02-analysis_data/analysis_data.parquet'))
-dataset <- data[,2:4]  
-target <- data[,5]
+data <- read_parquet(file = here::here("data/02-analysis_data/analysis_data.parquet"))
+dataset <- data[, 2:4]
+target <- data[, 5]
 
 n <- nrow(dataset)
 
@@ -20,21 +20,21 @@ predictions <- numeric(n)
 actuals <- target
 
 
-model <- read_rds(file = here::here('models/Mortality_Model.rds'))
+model <- read_rds(file = here::here("models/Mortality_Model.rds"))
 
 for (i in 1:n) {
   # Split data into training and testing sets
   train_data <- dataset[-i, ]
-  train_target <- target[-i,]
+  train_target <- target[-i, ]
   test_data <- dataset[i, , drop = FALSE]
-  
+
   # Predict for the left-out observation
   predictions[i] <- predict(model, newdata = test_data)
 }
 
 # Compute LOOCV R-squared
-ss_total <- sum((actuals - mean(unlist(actuals)))^2)  # Total Sum of Squares
-ss_residual <- sum((actuals - predictions)^2)  # Residual Sum of Squares
+ss_total <- sum((actuals - mean(unlist(actuals)))^2) # Total Sum of Squares
+ss_residual <- sum((actuals - predictions)^2) # Residual Sum of Squares
 r_squared <- 1 - (ss_residual / ss_total)
 
 # Print the LOOCV R-squared

### 2024-11-28 paper/paper.qmd
combine libraries

@@
 library(tidyverse)
+library(ggplot2)
+library(arrow)
+library(knitr)
+library(dplyr)
+library(modelsummary)
+library(rstanarm)
 ```
@@
 
-# Load necessary libraries
-library(ggplot2)
-library(arrow)
-
 # Read the data
@@
 
-# Load necessary libraries
-library(ggplot2)
-library(arrow)
-
 analysis_data <- read_parquet(file = (here::here("data/02-analysis_data/analysis_data.parquet")))
@@
 #| layout-ncol: 2
-
-# Load necessary libraries
-library(ggplot2)
-library(arrow)
-library(knitr)
 
@@
 #| layout-ncol: 2
-
-# Load necessary libraries
-library(ggplot2)
-library(arrow)
-library(knitr)
-library(dplyr)
 
@@
 #| tbl-cap: "Explanatory model of Mortality rate for each country based on food production index, DPT vaccine coverage, and health expense"
-
-#### Load libraries ####
-library(tidyverse)
-library(rstanarm)
-library(knitr)
-library(arrow)
-library(modelsummary)
 

### 2024-11-29 paper/paper.qmd
Discussion add reference to literature

@@
 
+The estimand of this study is the effect of three predictors—food production index, vaccination coverage (DPT vaccine percentage), and per-capita health expenditure (log-transformed)—on the logarithm of the under-five mortality rate for each country in 2022.
+This analysis seeks to quantify the extent to which these factors contribute to variations in mortality rates across nations.
+By addressing these relationships, the study aims to identify potential intervention points for reducing preventable child deaths and achieving progress toward the Sustainable Development Goals for child survival.
+
 The remainder of this paper is organized as follows.
@@
 Finally, @sec-diss discusses the implications of the findings, the limitations of the study, and recommendations for future research and policy interventions aimed at reducing child mortality globally.
-
-## Estimand
-The primary estimand of this study is the effect of three predictors—food production index, vaccination coverage (DPT vaccine percentage), and per-capita health expenditure (log-transformed)—on the logarithm of the under-five mortality rate for each country in 2022.
-This analysis seeks to quantify the extent to which these factors contribute to variations in mortality rates across nations.
-By addressing these relationships, the study aims to identify potential intervention points for reducing preventable child deaths and achieving progress toward the Sustainable Development Goals for child survival.
 
@@
 
-## Implications in reducing mortality rate
-The UN's Sustainable Development Goal (SDG) 3.2 aims to reduce the mortality rate of children under five to fewer than 25 per 1,000 live births by 2030 [@who2024].
-While substantial progress has been achieved, reaching this target within the next six years presents significant challenges.
+## Overview
+The United Nations’ Sustainable Development Goal (SDG) 3.2 aims to reduce the mortality rate of children under five to fewer than 25 per 1,000 live births by 2030 [@who2024].
+While substantial progress has been achieved globally, meeting this ambitious target within the remaining timeline of six years presents significant challenges.
 Although global trends indicate a decline in child mortality rates, the pace of this reduction has slowed since 2015 [@who2024].
-Moreover, as highlighted in @fig-mor-1, significant disparities persist between countries.
-Some nations have already surpassed the SDG target, with mortality rates well below the threshold, whereas others—primarily in low-income regions such as sub-Saharan Africa—experience rates four times higher than the target.
 
-To achieve the SDG targets, urgent and focused international collaboration is required. Our findings underscore two critical areas for intervention: vaccine coverage and healthcare spending.
+A stark disparity persists across countries, as illustrated in @fig-mor-1.
+While some nations, especially high-income ones, have already surpassed the SDG target, others—predominantly in low-income regions such as sub-Saharan Africa—experience child mortality rates that are four times higher than the global goal.
 
-Vaccination programs have a demonstrable and significant impact on reducing child mortality.
-As shown in @fig-DPT-1, the distribution of DPT vaccine coverage is skewed left, indicating that while most countries maintain adequate immunization rates, certain nations lag considerably behind.
-WHO’s analysis of DPT vaccine coverage highlights that low-performing countries are concentrated in Africa, the Eastern Mediterranean, and South-East Asia regions [@who_dtp3_coverage].
-These countries, predominantly developing nations, often face financial constraints that prevent them from funding vaccines for their newborn populations.
-Consequently, international aid is crucial to making DPT vaccines accessible and affordable in these regions.
+To achieve the SDG targets, concerted efforts are essential, particularly through targeted support from international organizations like the World Health Organization (WHO).
+However, this raises critical questions: Where should resources be allocated most effectively?
+How can countries with high child mortality rates achieve the greatest impact with limited resources?
 
-Healthcare spending presents a more complex challenge.
-It involves multiple factors, including a country’s economic capacity, infrastructure development, and access to healthcare services.
-Addressing these issues requires sustained, long-term efforts.
-Developed nations can play a pivotal role by providing financial support, sharing technological advancements, and facilitating knowledge transfer to strengthen healthcare systems in developing countries.
-Without addressing these systemic challenges, achieving equitable child survival rates will remain an unattainable goal.
+## Correlation V.S. Causality
+
+Numerous studies have explored the determinants of child mortality, consistently finding a negative correlation between mortality rates and macroeconomic indicators such as GDP [@ohare2013income] or education levels [@sartorius2014global].
+However, many of these factors are not direct causes of child mortality.
+
+For instance, during the Great Recession, the United States experienced a reduction in mortality rates due to decreased air pollution following widespread factory closures [@finkelstein2024lives].
+This demonstrates the importance of identifying underlying causes rather than focusing solely on broad economic indicators.
+In low-income countries, the primary drivers of high child mortality rates are often basic issues such as malnutrition and inadequate access to primary healthcare [@DOERR2022110860].
+Addressing these root causes is crucial to achieving meaningful and sustainable improvements.
+
+Our analysis identifies two factors with strong negative correlations to child mortality rates: DPT vaccine coverage and healthcare spending.
+However, it remains essential to determine whether these factors exert a direct influence or are proxies for other underlying variables.
+
+The DPT vaccine protects against three deadly diseases: diphtheria, pertussis (whooping cough), and tetanus (lockjaw).
+Administered in five doses starting at two months of age, this vaccine provides critical immunity for newborns [@cdc_dtap].
+
+In 1990, diphtheria, pertussis, and tetanus collectively caused approximately 160,000 deaths among children aged 1–4 years.
+By 2019, this number had decreased to 60,000, demonstrating the vaccine's impact [@gbd2019].
+Nonetheless, disparities in vaccination rates persist, leaving many children vulnerable to preventable diseases.
+Studies indicate that children receiving zero doses of vaccines are at significantly higher risk of mortality [@karlsson2024prevalence].
+Expanding vaccination coverage is a direct and cost-effective intervention to reduce child mortality and improve public health.
+
+Healthcare spending encompasses all expenditures on health-related goods and services within a given year, including both private and public spending [@worldbank_health_expenditure_us].
+Research suggests that public healthcare spending has the most pronounced effect on reducing child mortality [@ray2020health], primarily by increasing access to trained healthcare professionals and improving maternity and neonatal care.
+
+Public healthcare systems are often funded through tax revenues, which are tied to broader economic performance and governance structures [@ray2020health].
+While boosting overall tax revenues may be challenging, reallocating existing resources to prioritize public health can yield significant benefits.
+This approach is particularly relevant in countries where public spending on healthcare remains disproportionately low.
+
+## Conclusion
+
+Our study shows that reducing child mortality rate requires a dual focus on expanding vaccination coverage and optimizing healthcare spending.
+Vaccination programs, particularly for diseases like diphtheria, pertussis, and tetanus, represent a direct and impactful strategy for reducing child mortality.
+Simultaneously, enhancing public healthcare spending can address systemic issues such as inadequate maternal care and limited access to trained professionals.
+
+Efforts must be tailored to the specific needs and capacities of individual countries.
+International organizations can play a pivotal role by providing financial support, technical expertise, and policy guidance to countries with high child mortality rates.
+By prioritizing interventions with clear and measurable impacts, the global community can move closer to the goal of ensuring that no child dies from preventable causes.
 

### 2024-11-29 paper/paper.qmd
Through polish

@@
 
-Over the past 30 years, advancements in society and technology have significantly reduced global mortality rates, particularly for children under the age of five.
-The under-five mortality rate declined from 93 deaths per 1,000 live births in 1990 to 37 deaths per 1,000 live births in 2022 [@worldbank2024M].
+Over the past 30 years, advancements in society and technology have significantly reduced global mortality rates, particularly for children under the age of five (child mortality).
+The child mortality rate declined from 93 deaths per 1,000 live births in 1990 to 37 deaths per 1,000 live births in 2022 [@worldbank2024M].
 This progress represents millions of children who now have better chances of survival, marking a substantial achievement in global health.
 
-In this study, we examined the under-five mortality rate for each country in 2022, focusing on three key factors: food production index, DPT vaccine coverage, and per-capita health expenditure.
-These factors were chosen because they reflect critical components influencing child mortality.
-Infectious diseases, such as pneumonia, diarrhea, and malaria, remain the leading causes of under-five deaths, and these are directly linked to vaccination coverage and healthcare spending.
-Similarly, maternal health during childbirth is influenced by healthcare accessibility and quality, which is often tied to national health expenditures [@who2024childmortality].
-While food production does not directly address infectious disease or maternal health, it reflects a nation’s ability to meet nutritional needs, which is critical for preventing malnutrition—a major underlying cause of child deaths (add supporting explanation and citation).
+Despite this progress, the pace of reduction has slowed considerably since 2015, decreasing from an annual rate of 3.8% during the Millennium Development Goal (MDG) era (2000–2015) to 2.1% during the Sustainable Development Goal (SDG) era (2015–2022) [@unicef2022underfive].
+In 2022 alone, 4.9 million children under the age of five died, equating to approximately 13,400 deaths per day [@unicef2022underfive].
+Moreover, the distribution of mortality rates under five remains highly uneven.
+Developed countries such as Canada and the United States report mortality rates as low as 10 per 1,000 live births, while developing nations—particularly in sub-Saharan Africa—experience rates exceeding 100 per 1,000 live births [@worldbank].
+This stark disparity underscores the urgent need for targeted interventions in high-mortality regions.
 
+Infectious diseases, such as pneumonia, diarrhea, and malaria, remain the leading causes of under-five deaths, closely tied to vaccination coverage and healthcare spending.
+Maternal health is also influenced by healthcare accessibility and quality, which often correlates with national health expenditures [@who2024childmortality].
+While food production does not directly address these issues, it reflects a nation’s ability to meet nutritional needs, a critical factor in preventing malnutrition—a major contributor to child mortality.
+
+In this study, we examined the child mortality rate for each country in 2022, focusing on three key factors: food production index, DPT vaccine coverage, and per-capita health expenditure.
 Our findings reveal that increasing vaccine coverage and per-capita health expenditures significantly reduce under-five mortality rates.
-However, the food production index does not exhibit a direct impact, suggesting that its role is mediated by other factors such as distribution systems and food quality.
-Despite progress in reducing global mortality, the pace of reduction has slowed considerably since 2015, decreasing from an annual rate of 3.8% during the Millennium Development Goal (MDG) era (2000–2015) to 2.1% during the Sustainable Development Goal (SDG) era (2015–2022) [@unicef2022underfive].
-In 2022 alone, 4.9 million children under the age of five died, equating to approximately 13,400 deaths per day [@unicef2022underfive].
-
-Furthermore, the distribution of mortality rates under five remains highly uneven.
-Developed countries such as Canada and the United States report mortality rates as low as 10 per 1,000 live births while developing nations—particularly in sub-Saharan Africa—experience rates exceeding 100 per 1,000 live births (@worldbank).
-This disparity highlights the urgent need for targeted interventions in high-mortality regions.
-The factors studied in this paper, such as vaccination and healthcare spending, are actionable and can be effectively improved in these regions through foreign aid and international collaboration.
+However, the food production index does not exhibit a direct impact, suggesting its role is mediated by factors such as food quality and distribution systems.
 
@@
 
-For this analysis, we combined four data sets all together into one.
-All these four datasets are downloaded through **Worldbank Open Data** API [@worldbank].
+For this analysis, we used four datasets: child mortality rate, food production index, DPT vaccine coverage rate, and health expense per capita.
+All four datasets are downloaded through **Worldbank Open Data** API [@worldbank].
 We employed **R** [@citeR], a coding platform to download, clean, and conduct statistical analysis.
@@
 
-The analysis in this study relies on four data sets, each collected and prepared by various UN-affiliated organizations.
+The datasets of this analysis are collected and prepared by various UN-affiliated organizations.
 Below, we provide a brief overview of the measurement and transformation processes for these data sets.
@@
 The primary data source is each country's vital registration system.
-In countries such system is lacking, household surveys are conducted and then processed using a statistical model developed collaboratively by UN agencies and academic institutions to derive robust mortality rate estimates [@worldbank2024M].
+In countries such system is lacking, household surveys are conducted and then processed using a statistical model developed collaboratively by UN agencies and academic institutions to derive an estimate for mortality rate [@worldbank2024M].
 
@@
 
-The response variable in this study is the under-five mortality rate, measured as the number of deaths of children under five years old per 1,000 live births.
+The response variable in this study is the child mortality rate in each country in year 2022, measured as the number of deaths of children under five years old per 1,000 live births in year 2022.
 
@@
 Due to extreme values and significant right skewness, a log transformation was applied to stabilize variance and improve normality.
-The transformed data range from 2 to 5, with a peak near 3 (@fig-mor-2).
+The transformed data range from 2 to 5, with a peak near 3, and the shape is approximately normal.
+(@fig-mor-2).
 
@@
 **Food Production Index:** 
-The Food Production Index measures the relative level of agricultural production for each country, using the 2014–2016 baseline as a reference point [@worldbank].
+The Food Production Index measures the relative level of agricultural production for each country in year 2022, using the 2014–2016 baseline as a reference point [@worldbank].
 This index is price-weighted and reflects agricultural productivity across countries.
 As illustrated in @fig-food-1, the index ranges from 60 to 180, with a mean of around 110 and an approximately normal distribution.
-Scatter plot analysis in @fig-food-2 reveals a slightly positive linear relationship between the Food Production Index and the log-transformed mortality rate.
-This suggests that as agricultural production increases, mortality rates tend to rise modestly, a pattern that warrants further exploration in @sec-model.
+Scatter plot between food production index and log of child mortality in @fig-food-2 shows a slightly positive linear relationship.This relationship is examined in greater detail in @sec-results.
 ```{r}
@@
 The histogram on the left shows a approximately normal distrubution.
-The scatterplot on the right indicates there is a slight positive relationship between Food Production Index and Mortality rate
+The scatterplot on the right indicates there is a slight positive relationship between food production index and mortality rate
 #| fig-subcap: ["Counts of Food production index for each country, measured as a basis of 2014-1016", "Relationship of each countries' Food production index VS. Log of Mortality rate"]
@@
 **Current Health Expenditure:**
-The Current Health Expenditure per Capita captures health spending in US dollars, calculated under the System of Health Accounts (SHA 2011) framework.
+The Current Health Expenditure per Capita captures health spending in US dollars for each individual in year 2021, calculated under the System of Health Accounts (SHA 2011) framework.
 The raw data range from $0 to $12,000 and exhibit extreme right skewness, as shown in @fig-health-1.
@@
 Analysis of the relationship between log-transformed health expenditure and log mortality rate (@fig-health-3) demonstrates a significant negative linear trend, indicating that higher health expenditures are associated with lower mortality rates.
-This relationship is examined in greater detail in @sec-model.
+This relationship is examined in greater detail in @sec-results.
 ```{r}
@@
 **DPT Vaccine Coverage:**
-DPT vaccine coverage rate reflects the percentage of children aged 12–23 months who received vaccinations against diphtheria, pertussis, and tetanus.
+DPT vaccine coverage rate reflects the percentage of children aged 12–23 months who received vaccinations against diphtheria, pertussis, and tetanus in year 2022.
 Coverage ranges from 30% to 100%, with a left-skewed distribution, as depicted in @fig-DPT-1.
@@
 
-Additionally, some smaller countries lacked complete data and were excluded from the analysis. The implications of these omissions are addressed in @sec-results.
+Additionally, some countries' data are not collected and were excluded from the analysis. The implications of these omissions and time inconsistancy are addressed in @sec-lim.
 
@@
 By analyzing these factors, we seek to identify opportunities to improve child health outcomes, particularly in countries with high mortality rates.
-Supporting background information and diagnostics are provided in [Appendix -@sec-model-details].
+Supporting model diagnostics are provided in [Appendix -@sec-model-details].
 
@@
 
-We implement this model in R [@citeR] using the rstanarm package [@rstanarm].
+We implement this model in **R** [@citeR] using the **rstanarm**  [@rstanarm] and **arrow** [@arrow] package.
 
@@
 Default Bayesian priors were used for the predictors, as they provide a robust starting point.
-For the intercept term, a \mbox{Normal}(2.5, 1) prior was chosen, reflecting the observed distribution of the response variable (see @sec-data).
+For the intercept term, a \mbox{Normal}(2.5, 1) prior was chosen, reflecting the observed distribution of the response variable (@fig-mor-1).
 
@@
 
-2. Potential Collinearity: While each predictor shows an individual linear relationship with the response variable, interdependencies between predictors may affect model accuracy.
+2. Potential Collinearity: While each predictor shows an individual linear relationship with the response variable, dependencies between predictors may affect model accuracy.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+# Model outcome
+
+Our model results are summarized in @tbl-modelresults .We will discuss them one by one
 
@@
 
+# Model Validation
+The posterior and prior predictive checks are presented in @fig-ppcheckandposteriorvsprior-1.
+We observe an excellent fit between the posterior distribution from our model and the actual child mortality data, indicating that the model effectively captures the underlying trends in child mortality rates.
+In @fig-ppcheckandposteriorvsprior-2, we compare the posterior distribution with the prior.
+No significant deviations are observed, suggesting alignment between our prior beliefs and the posterior distribution.
+
+@fig-stanareyouokay-1 displays the trace plot check, where no irregularities or concerning patterns are detected.
+Finally, the R-hat plot (@fig-stanareyouokay-2) shows all values close to one, confirming the convergence of the Markov chains.
+
 # Discussion {#sec-diss}
@@
 In low-income countries, the primary drivers of high child mortality rates are often basic issues such as malnutrition and inadequate access to primary healthcare [@DOERR2022110860].
-Addressing these root causes is crucial to achieving meaningful and sustainable improvements.
+Addressing such root causes is crucial to achieving meaningful and sustainable improvements.
 
@@
 Our study shows that reducing child mortality rate requires a dual focus on expanding vaccination coverage and optimizing healthcare spending.
-Vaccination programs, particularly for diseases like diphtheria, pertussis, and tetanus, represent a direct and impactful strategy for reducing child mortality.
+Vaccination programs, particularly for diseases like diphtheria, pertussis, and tetanus, represent a direct strategy for reducing child mortality.
 Simultaneously, enhancing public healthcare spending can address systemic issues such as inadequate maternal care and limited access to trained professionals.
 
-Efforts must be tailored to the specific needs and capacities of individual countries.
+Oh the other hand, efforts must be tailored to the specific needs and capacities of individual countries.
 International organizations can play a pivotal role by providing financial support, technical expertise, and policy guidance to countries with high child mortality rates.
@@
 
-## Limitation and next steps
-This study has several limitations that should be considered in future research.
-First, as noted in @sec-data-missing, data from several countries is unavailable.
-While these nations have minimal populations, excluding them introduces potential biases and raises ethical concerns by overlooking their unique challenges.
-Second, the most recent data used in this study is from 2022.
-Access to more current data sets would enable a more accurate and relevant analysis, reflecting recent developments in healthcare and child mortality trends.
+## Limitation and next steps {#sec-lim}
+This study has several limitations that warrant consideration in future research.
+First, as discussed in @sec-data-missing, data from several countries is unavailable.
+While these nations typically have smaller populations, their exclusion may introduce biases and raises ethical concerns by neglecting their unique challenges.
+Second, the most recent data used in this study is from 2022, and there are temporal inconsistencies among the predictors.
+This is particularly important because both the predictors and the response variable are time-sensitive.
+Access to more current and synchronized datasets would enable a more accurate and relevant analysis, reflecting recent developments in healthcare and child mortality trends.
 
-Additionally, while this study focused on three primary predictors, the dataset includes other indicators that could potentially enrich the analysis.
-Unfortunately, many of these indicators suffer from missing or outdated values, limiting their utility.
-Future research should prioritize data completeness and timeliness to improve model robustness and reliability.
+Additionally, while this study focused on three primary predictors, the dataset contains other indicators that could potentially enhance the analysis.
+However, many of these indicators are affected by missing or outdated values, limiting their utility.
+Future research should prioritize addressing data completeness and timeliness to improve model robustness and reliability.
 
-Finally, advanced modeling techniques and the inclusion of additional predictors could yield deeper insights.
-Future studies should explore innovative models to better capture complex relationships and interactions, offering more nuanced recommendations for policy interventions.
+Finally, incorporating advanced modeling techniques and additional predictors could provide deeper insights.
+Future studies should explore innovative models to better capture complex relationships and interactions, enabling more nuanced and actionable recommendations for policy interventions.
 

### 2024-11-29 paper/paper.qmd
Final submit

@@
 abstract: "Child mortality has significantly decreased over the past three decades, but progress remains uneven across regions.
-This study examines the under-five mortality rates of 2022, focusing on the effects of vaccination coverage, health expenditure per capita, and food production indices.
-Using a Bayesian modeling approach, we found that increased vaccination coverage and healthcare spending significantly reduce child mortality, while food production had no direct impact.
-These findings underscore the critical need for targeted interventions in high-mortality regions, emphasizing the importance of vaccination programs and healthcare funding in achieving global child survival goals."
+This study examines the under-five mortality rates of 2022 for each country, focusing on the effects of DPT vaccination coverage, health expenditure per capita, and food production indices.
+Using a Bayesian multi-linear modeling approach, we found that increased vaccination coverage and healthcare spending significantly reduce child mortality, while food production had no direct impact.
+These findings underscore the need for targeted interventions in high-mortality regions, suggesting the importance of vaccination programs and healthcare funding in achieving global child survival goals."
 format: pdf
@@
 
-## Overview Check packages
+## Overview
 
@@
 All four datasets are downloaded through **Worldbank Open Data** API [@worldbank].
+World Bank Open Data platform is a comprehensive resource offering free and publicly accessible development data from nearly every country.
+This platform provides an extensive range of indicators across sectors such as health, agriculture, and education, enabling researchers to access globally standardized data for comparative analysis.
+The platform ensures transparency and usability by offering tools for visualization, data export, and API integration, making it a reliable and versatile resource for exploring global trends.
+
 We employed **R** [@citeR], a coding platform to download, clean, and conduct statistical analysis.
@@
 
+## Model Validation
+The posterior and prior predictive checks are presented in @fig-ppcheckandposteriorvsprior-1.
+We observe an excellent fit between the posterior distribution from our model and the actual child mortality data, indicating that the model effectively captures the underlying trends in child mortality rates.
+In @fig-ppcheckandposteriorvsprior-2, we compare the posterior distribution with the prior.
+No significant deviations are observed, suggesting alignment between our prior beliefs and the posterior distribution.
+
+@fig-stanareyouokay-1 displays the trace plot check, where no irregularities or concerning patterns are detected.
+Finally, the R-hat plot (@fig-stanareyouokay-2) shows all values close to one, confirming the convergence of the Markov chains.
+Detailed information could be foun in @sec-model-details.
+
 # Results {#sec-results}
 
-# Model outcome
+## Model outcome
 
@@
 
-## Intercept:
+### Intercept:
 The intercept estimate of 6.643 represents the expected logarithm of the under-five mortality rate when all predictors are held constant at their reference or baseline values (e.g., average food production, average vaccine coverage, and average health expenditure).
 
-## Food Production Index:
+### Food Production Index:
 The coefficient for the food production index is 0.003, indicating a very slight positive relationship between food production and the under-five mortality rate.
@@
 
-## Vaccine Coverage:
+### Vaccine Coverage:
 The coefficient for vaccine coverage is -0.013, meaning that for every 1 percentage point increase in vaccine coverage, the logarithm of the under-five mortality rate decreases by 1.3%, holding other variables constant.
 
-## Health Expenditure (Log-Transformed):
+### Health Expenditure (Log-Transformed):
 The coefficient for the log-transformed health expenditure is -0.529, indicating that a 1% increase in health expenditure per capita is associated with a 0.529 unit decrease in the log of the under-five mortality rate, holding all other predictors constant.
@@
 ```
-
-# Model Validation
-The posterior and prior predictive checks are presented in @fig-ppcheckandposteriorvsprior-1.
-We observe an excellent fit between the posterior distribution from our model and the actual child mortality data, indicating that the model effectively captures the underlying trends in child mortality rates.
-In @fig-ppcheckandposteriorvsprior-2, we compare the posterior distribution with the prior.
-No significant deviations are observed, suggesting alignment between our prior beliefs and the posterior distribution.
-
-@fig-stanareyouokay-1 displays the trace plot check, where no irregularities or concerning patterns are detected.
-Finally, the R-hat plot (@fig-stanareyouokay-2) shows all values close to one, confirming the convergence of the Markov chains.
 

### 2024-11-29 paper/paper.qmd
data citation

@@
 Moreover, the distribution of mortality rates under five remains highly uneven.
-Developed countries such as Canada and the United States report mortality rates as low as 10 per 1,000 live births, while developing nations—particularly in sub-Saharan Africa—experience rates exceeding 100 per 1,000 live births [@worldbank].
+Developed countries such as Canada and the United States report mortality rates as low as 10 per 1,000 live births, while developing nations—particularly in sub-Saharan Africa—experience rates exceeding 100 per 1,000 live births [@worldbank2024M].
 This stark disparity underscores the urgent need for targeted interventions in high-mortality regions.
@@
 For this analysis, we used four datasets: child mortality rate, food production index, DPT vaccine coverage rate, and health expense per capita.
-All four datasets are downloaded through **Worldbank Open Data** API [@worldbank].
-World Bank Open Data platform is a comprehensive resource offering free and publicly accessible development data from nearly every country.
+All four datasets are downloaded from **Worldbank Open Database** [@worldbank_data] through its API [@worldbank].
+World Bank Open Database is a comprehensive resource offering free and publicly accessible development data from nearly every country.
 This platform provides an extensive range of indicators across sectors such as health, agriculture, and education, enabling researchers to access globally standardized data for comparative analysis.
@@
 We employed **R** [@citeR], a coding platform to download, clean, and conduct statistical analysis.
-Besides, we also utilized R packages **tidyverse** [@tidy], **rstanarm** [@rstanarm], **ggplot2** [@gg], **knitr** [@knitr], **arrow** [@arrow], **here** [@here], **modelsummary** [@ms], and **dpylr** [@dplyr].
+Besides, we also utilized R packages **tidyverse** [@tidy], **rstanarm** [@rstanarm], **ggplot2** [@gg], **knitr** [@knitr], **arrow** [@arrow], **here** [@here], **modelsummary** [@ms], **testthat** [@testthat] and **dpylr** [@dplyr].
 The paper is outlined in Git Hub using the starter folder provided in **Telling Stories With Data** [@tellingstories].

### 2024-12-05 paper/paper.qmd
The GitHub link display issue in the paper has been corrected. This update was made following a valuable suggestion from University of Totonto Prof. Rohan Alexander (https://rohanalexander.com/). We appreciate his insightful feedback.

@@
   - Junbo Li
-thanks: "Code and data are available at: [https://github.com/Junbo345/Mortality_analysis]."
+thanks: "Code and data are available at: [https://github.com/Junbo345/Mortality_analysis](https://github.com/Junbo345/Mortality_analysis)."
 date: today

### 2024-12-10 paper/paper.qmd
grammar error change

@@
 Moreover, the distribution of mortality rates under five remains highly uneven.
-Developed countries such as Canada and the United States report mortality rates as low as 10 per 1,000 live births, while developing nations—particularly in sub-Saharan Africa—experience rates exceeding 100 per 1,000 live births [@worldbank2024M].
+Developed countries such as Canada and the United States report mortality rates as low as 10 per 1,000 live births while developing nations—particularly in sub-Saharan Africa—experience rates exceeding 100 per 1,000 live births [@worldbank2024M].
 This stark disparity underscores the urgent need for targeted interventions in high-mortality regions.
@@
 The primary data source is each country's vital registration system.
-In countries such system is lacking, household surveys are conducted and then processed using a statistical model developed collaboratively by UN agencies and academic institutions to derive an estimate for mortality rate [@worldbank2024M].
+In countries where such a system is lacking, household surveys are conducted and then processed using a statistical model developed collaboratively by UN agencies and academic institutions to derive an estimate of the mortality rate [@worldbank2024M].
 
@@
 **Food Production Index**: This index, prepared by the Food and Agriculture Organization (FAO) of the United Nations, measures aggregate food production relative to the 2014–2016 baseline period.
-It is calculated using price-weighted data on agricultural products for each country yearly.
+It is calculated yearly using price-weighted data on agricultural products for each country.
 [@worldbank_foodindex2024].
@@
 **Health Expenditure per Capita**: Data on health expenditure per capita are produced by the World Health Organization (WHO) following the System of Health Accounts 2011 (SHA 2011) framework.
-This framework tracks all health-related spending in a country over a defined time period [@worldbank_health_expenditure_us].
+This framework tracks all health-related spending in a country over a defined period [@worldbank_health_expenditure_us].
 
@@
 
-The response variable in this study is the child mortality rate in each country in year 2022, measured as the number of deaths of children under five years old per 1,000 live births in year 2022.
+The response variable in this study is the child mortality rate in each country in 2022, measured as the number of deaths of children under five years old per 1,000 live births in 2022.
 
@@
 #| label: fig-mor
-#| fig-cap: Data Analysis for Mortality Rate for children under five of different countries.
+#| fig-cap: Data Analysis for Mortality Rate for children under five in different countries.
 The histogram on the top left shows that the original distribution is skewed to the right.
 The histogram on the top right is the distribution after log transformation and is approximately normal.
-#| fig-subcap: ["Dsitribution of Mortality rate before log tranformation", "Distribution of Mortality rate after log transformation"]
+#| fig-subcap: ["Distribution of Mortality rate before log transformation", "Distribution of Mortality rate after log transformation"]
 #| echo: false
@@
 **Food Production Index:** 
-The Food Production Index measures the relative level of agricultural production for each country in year 2022, using the 2014–2016 baseline as a reference point [@worldbank].
+The Food Production Index measures the relative level of agricultural production for each country in 2022, using the 2014–2016 baseline as a reference point [@worldbank].
 This index is price-weighted and reflects agricultural productivity across countries.
 As illustrated in @fig-food-1, the index ranges from 60 to 180, with a mean of around 110 and an approximately normal distribution.
-Scatter plot between food production index and log of child mortality in @fig-food-2 shows a slightly positive linear relationship.This relationship is examined in greater detail in @sec-results.
+The scatter plot between the food production index and log of child mortality in @fig-food-2 shows a slightly positive linear relationship.
+This relationship is examined in greater detail in @sec-results.
 ```{r}
@@
 #| fig-cap: Data Analysis of Predictor Food Production Index.
-The histogram on the left shows a approximately normal distrubution.
-The scatterplot on the right indicates there is a slight positive relationship between food production index and mortality rate
+The histogram on the left shows an approximately normal distribution.
+The scatterplot on the right indicates there is a slight positive relationship between food production index and mortality rate.
 #| fig-subcap: ["Counts of Food production index for each country, measured as a basis of 2014-1016", "Relationship of each countries' Food production index VS. Log of Mortality rate"]
@@
 **Current Health Expenditure:**
-The Current Health Expenditure per Capita captures health spending in US dollars for each individual in year 2021, calculated under the System of Health Accounts (SHA 2011) framework.
+The Current Health Expenditure per Capita captures health spending in US dollars for each individual in 2021, calculated under the System of Health Accounts (SHA 2011) framework.
 The raw data range from $0 to $12,000 and exhibit extreme right skewness, as shown in @fig-health-1.
@@
 The histogram on the top right is the distribution after log transformation and is approximately normal.The scatterplot on the bottom left indicates there is a negative relationship between log of health expenses and Mortality rate.
-#| fig-subcap: ["Counts of each countries' Health Expenditure Per Capita, Measured in US$.
-Before log transformation", "Counts of each countries' Health Expenditure Per Capita, Measured in US$.
-After log transformation", "Relationship of each countries' Health Expenditure Per Capita, Measured in US$ VS.
+#| fig-subcap: ["Counts of Health Expenditure Per Capita for each country (measured in US$) before log transformation", "Counts of each country's Health Expenditure Per Capita, (measured in US$) after log transformation", "Relationship of each countries' Health Expenditure Per Capita, Measured in US$ VS.
 Log of Mortality"]
@@
 **DPT Vaccine Coverage:**
-DPT vaccine coverage rate reflects the percentage of children aged 12–23 months who received vaccinations against diphtheria, pertussis, and tetanus in year 2022.
+DPT vaccine coverage rate reflects the percentage of children aged 12–23 months who received vaccinations against diphtheria, pertussis, and tetanus in 2022.
 Coverage ranges from 30% to 100%, with a left-skewed distribution, as depicted in @fig-DPT-1.
@@
 #| label: fig-DPT
-#| fig-cap: Data Analysis of DPT vaccine coverage.
-The histogram on the left shows that the distribution DPT vaccine coverage of is skewed to the left.
+#| fig-cap: Data Analysis of DPT Vaccine Coverage.
+The histogram on the left shows that the distribution of DPT vaccine coverage is skewed to the left.
 The scatterplot on the right indicates there is a negative relationship between DPT vaccine coverage and Mortality rate.
-#| fig-subcap: ["Counts of each countries' DPT vaccine percentage", "Relationship of each countries' DPT Vaccine Percentage VS. Log of Mortality"]
+#| fig-subcap: ["Counts of each country's DPT vaccine percentage", "Relationship of each country's DPT Vaccine Percentage VS. Log of Mortality"]
 #| echo: false
@@
 ```
-## Missing Data and Time Inconsistancy {#sec-data-missing}
+## Missing Data and Time Inconsistency {#sec-data-missing}
 The health expenditure variable was collected in 2021, while other variables reflect 2022 data.
@@
 
-Additionally, some countries' data are not collected and were excluded from the analysis. The implications of these omissions and time inconsistancy are addressed in @sec-lim.
+Additionally, some countries' data were not collected and were excluded from the analysis. The implications of these omissions and time inconsistency are addressed in @sec-lim.
 
@@
 
-Let $y_i$ be the logarithm of under-five mortality rate for nation $i$. We define the following predictors:
+Let $y_i$ be the logarithm of the under-five mortality rate for nation $i$. We define the following predictors:
 
@@
 Finally, the R-hat plot (@fig-stanareyouokay-2) shows all values close to one, confirming the convergence of the Markov chains.
-Detailed information could be foun in @sec-model-details.
+Detailed information could be found in @sec-model-details.
 
@@
 
-Our model results are summarized in @tbl-modelresults .We will discuss them one by one
+Our model results are summarized in @tbl-modelresults. We will discuss them one by one.
 
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory model of Mortality rate for each country based on food production index, DPT vaccine coverage, and health expense"
+#| tbl-cap: "Explanatory model of Mortality rate for each country based on food production index, DPT vaccine coverage, and health expense."
 
@@
 The United Nations’ Sustainable Development Goal (SDG) 3.2 aims to reduce the mortality rate of children under five to fewer than 25 per 1,000 live births by 2030 [@who2024].
-While substantial progress has been achieved globally, meeting this ambitious target within the remaining timeline of six years presents significant challenges.
+While substantial progress has been achieved globally, meeting this ambitious target within the remaining six years presents significant challenges.
 Although global trends indicate a decline in child mortality rates, the pace of this reduction has slowed since 2015 [@who2024].
@@
 
-Our study shows that reducing child mortality rate requires a dual focus on expanding vaccination coverage and optimizing healthcare spending.
+Our study shows that reducing the child mortality rate requires a dual focus on expanding vaccination coverage and optimizing healthcare spending.
 Vaccination programs, particularly for diseases like diphtheria, pertussis, and tetanus, represent a direct strategy for reducing child mortality.
@@
 
-Oh the other hand, efforts must be tailored to the specific needs and capacities of individual countries.
+On the other hand, efforts must be tailored to the specific needs and capacities of individual countries.
 International organizations can play a pivotal role by providing financial support, technical expertise, and policy guidance to countries with high child mortality rates.
@@
 First, as discussed in @sec-data-missing, data from several countries is unavailable.
-While these nations typically have smaller populations, their exclusion may introduce biases and raises ethical concerns by neglecting their unique challenges.
+While these nations typically have smaller populations, their exclusion may introduce biases and raise ethical concerns by neglecting their unique challenges.
 Second, the most recent data used in this study is from 2022, and there are temporal inconsistencies among the predictors.

### 2025-04-14 paper/paper.qmd
Intro

@@
 # Introduction
+
+Strong lensing occurs when a massive foreground object, such as a galaxy cluster, bends light from a more distant source, producing distinctive arcs or rings in images.
+Despite their significance for cosmology, these events are extremely rare, making up only a small fraction of astronomical observations.
+Traditional supervised learning methods require large quantities of labeled data, which is infeasible in this context.
+The primary challenge, therefore, is to develop a system capable of identifying potential lensing candidates with minimal labeled data.
 

### 2025-04-14 paper/paper.qmd
new

@@
 Despite their significance for cosmology, these events are extremely rare, making up only a small fraction of astronomical observations.
-Traditional supervised learning methods require large quantities of labeled data, which is infeasible in this context.
+Traditional supervised learning methods require large quantities of labeled data, which is unfeasible in this context.
 The primary challenge, therefore, is to develop a system capable of identifying potential lensing candidates with minimal labeled data.
+
+Thus we will use the state of art active learning strategy, BADGE, which aim to selects the most informative data points for human annotation and thus would improve training efficiency of the Machine Learning model [@cite].
+This strategy have already had promising outcomes in many lacking labeled classification tasks and is of the most effective ways in active learning [@cite].
+
+However, we are still facing another problem which is the imbalanced data.
+the strong lensing ones in our dataset only consists 8% of the total data.
+And thus will make the evaluation and traing very trick.
+So we will incorporate better evaluating metrics and training loss.
+
+The remainder of this paper is organized as follows.
+@sec-method describes possible losses functions we incorporate into our model.
+@sec-metric introduces the new metrics we incorporated @sec-results presents the results of new losses and evaluation metric.
+Finally, @sec-diss discusses the implications of the findings, the limitations of the study, and recommendations for future research in finding strong lensing data.
 

### 2025-04-14 paper/paper.qmd
Method

@@
 Finally, @sec-diss discusses the implications of the findings, the limitations of the study, and recommendations for future research in finding strong lensing data.
+
+# Method
+
+To test the new metric and loss functions, we use the available strong lensing data for regression.
+We concatenate the features extracted from the CNN and contatenate them with the expert labels for regression.
+The result data has a total of ~ and ~ strong lenseing ones.
+
+## New loss function
+
+Here we will investigate two loss functions, ranknet loss and focal loss.
 

### 2025-04-14 paper/paper.qmd
Method

@@
 Here we will investigate two loss functions, ranknet loss and focal loss.
+Ranknet loss is sepcified as follow: Given training data $x_1$, $x_2$, ....
+, $x_n$, and their rankings, $y_1$, $y_2$, ..., $y_n$, we will investigate all combinations of ordered pairs ($x_i$, $x_j$).
+The total loss is:
+$$
+L = \sum_{i < j} - y_{ij} \log P(i > j) - (1 - y_{ij}) \log (1 - P(i > j))
+$$
+Where $y_ij$ is 1 if $x_i$ is ranked higher than $y_j$, 0 if $x_i$ is ranked lower than $y_j$ (here we will assume no $x_i$ can have the same ranking as $x_j$ if i is not equal to j).
+$ P(i > j) = \frac{1}{1 + e^{-(s(x_i) - s(x_j))}} $ is the probability that $ x_i $ is ranked higher than $ x_j $.
+s() is the rankning score assigned to i, higher s higher ranking.
 

### 2025-04-14 paper/paper.qmd
Method

@@
 
-# Method
+# Method {#sec-method}
 
@@
 Where $y_ij$ is 1 if $x_i$ is ranked higher than $y_j$, 0 if $x_i$ is ranked lower than $y_j$ (here we will assume no $x_i$ can have the same ranking as $x_j$ if i is not equal to j).
-$ P(i > j) = \frac{1}{1 + e^{-(s(x_i) - s(x_j))}} $ is the probability that $ x_i $ is ranked higher than $ x_j $.
-s() is the rankning score assigned to i, higher s higher ranking.
+$P(i > j) = \frac{1}{1 + e^{-(s(x_i) - s(x_j))}}$ is the probability that $x_i$ is ranked higher than $x_j$.
+s() is the rankning score assigned to i, higher ranking cores means higher ranking.
+This loss aims to peanalize the pairs of data $x_i$ and $x_j$ where $x_i$ should be ranked higher than $x_j$ but instead $x_i$ is ranked lower by the model for $x_j$.
 

### 2025-04-15 paper/paper.qmd
Result

@@
 This loss aims to peanalize the pairs of data $x_i$ and $x_j$ where $x_i$ should be ranked higher than $x_j$ but instead $x_i$ is ranked lower by the model for $x_j$.
+
+The focal loss is specified as follows: 
+$$
+FL(p_t) = -\alpha_t (1-p_t)^\gamma \log(p_t)
+$$
+Here $p_t$ is the model's predicted probility if the true label is strong lensing, one minus the predicted probbility if the true label is not strong lenseing.
+$a_t$ is 0.75 if the true lavel is strong lensing and 0.25 if the true label is not strong lensing.
+The aim of this loss function is to avoid easy labels and penalize the model most if it incorrectly label a strong lensing one as non-lensing.
+
+## New evaluating matrics
+
+Despite the fact that we need new loss functions, we also need new evaluation metrics to better evaluate the performance of model on biased labels.
+As said, the labels of strong is very rare relative to non lensing.
+So if we use the default prediction accuracy then by just predicting everything by strong lensing the model will get a very decent performance while did not contribute much to the final goal of our research purpose.
+
+Thus we will propose a practical and better to understand metric, which is the # of correct labels in the top 10% of the highest ranked data.
+This is more to common sense and align the final purpose of our research.
+
+# Results
 

### 2025-04-15 paper/paper.qmd
Result

@@
 Ranknet loss is sepcified as follow: Given training data $x_1$, $x_2$, ....
-, $x_n$, and their rankings, $y_1$, $y_2$, ..., $y_n$, we will investigate all combinations of ordered pairs ($x_i$, $x_j$).
+, $x_n$, and their rankings, $y_1$, $y_2$, ..., $y_n$, we will investigate all combinations of ordered pairs ($x_i$, $x_j$) where their labels $y_i$ and $y_j$ are different.
 The total loss is:
@@
 # Results
+
+To test the two loss functions useing the new metric we build a one layer neuronetwork with 40 input neurons, 128 hidden neurons and one output which is the score of the input data in pytorch.
+We split the dataset into training and testing parts with the ratio of 7:3.
+To gain efficiency in training, we split the training data into minibatches of size 128 and after finetuning set the training epohs be 100 to avoid overtraining.
+We also used a neuronet work with crossentropy loss as a conterpart.
+We ran 10 independent groups and the result is the following:
 

### 2025-04-15 paper/paper.qmd
Result

@@
 We ran 10 independent groups and the result is the following:
+
+```{r}
+#| label: fig-mor1
+#| fig-cap: "Performance of two new proposed loss functions compared to crossentropy loss.
+They are evaluated based on number of correct labels in the top 100 labels predicted by the neuronetwork, roughly 10% of all the testing data.
+We see that rank loss does not outperform crossentropy loss while focal loss did but there is no promising beatens since the error bars contain each other's mean."
+#| echo: false
+#| warning: false
+#| message: false
+# Load required libraries
+library(ggplot2)
+library(dplyr)
+
+# Define data
+mlp <- c(36, 34, 38, 40, 44, 39, 42, 37, 40, 42)
+ranknet <- c(34, 35, 33, 40, 44, 40, 44, 39, 42, 37)
+focal <- c(35, 38, 39, 41, 41, 40, 50, 40, 43, 41)
+
+# Create a data frame
+df <- data.frame(
+  Model = rep(c("MLP", "RankNet", "Focal Loss"), each = 10),
+  Correct = c(mlp, ranknet, focal)
+)
+
+# Summarize: compute means and standard deviations
+summary_df <- df %>%
+  group_by(Model) %>%
+  summarise(
+    Mean = mean(Correct),
+    SD = sd(Correct)
+  )
+
+# Plot
+ggplot(summary_df, aes(x = Model, y = Mean, fill = Model)) +
+  geom_bar(stat = "identity", width = 0.6, color = "black") +
+  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.2) +
+  geom_text(aes(label = round(Mean, 1)), vjust = -0.5, size = 5) +
+  labs(
+    title = "Performance of MLP, RankNet, and Focal Loss Models Across 10 Splits",
+    y = "Correct Predictions in Top 100",
+    x = ""
+  ) +
+  theme_minimal() +
+  theme(
+    legend.position = "none",
+    text = element_text(size = 14)
+  )
+
+```
 

### 2025-04-15 paper/paper.qmd
First Draft

@@
 ```
+
+@fig-mor1 show the performance of two new proposed loss functions compared to crossentropy loss.
+They are evaluated based on average number of correct labels in the top 100 labels predicted by the neuronetwork, roughly 10% of all the testing data.
+The performace of three losses are quite th same, centered around 40 labels.
+We see that rank loss does not outperform crossentropy loss while focal loss did but there is no promising beatens since the error bars contain each other's mean.
+
+# Dsicussion
+
+## Conslusion and limitation
+
+To our experiment, we see that the Focal loss would be the best choice, but in reality it will requires more computational complexity in training and espically in BADGE where we need to find the gredient wrt the second last layer of the neuronet work.
+With cross entropy, the gredient has a nice closed form, which is ...
+However, using focal loss the gredient is much mroe computational cost than cross entropy becasue of the $a_t$ and $\gamma$ term added.
+Thus we still need to either finetuning the hyperparameters of the loss to improve its performance.
+As for ranknet loss, since we need the labels of two diff data be diff, and the case is very rare in each batch and thus will result some training batches yeild non pairs.
+
+## Future work
+
+Although we encountered certain limitations during the project, our findings open the door to several promising avenues for future work.
+These include fine-tuning the CNN feature extractor that supplies input to our model, enhancing the loss function or exploring alternative ranking approaches, and refining the evaluation criteria to better align with real-world requirements.
+Additionally, incorporating active learning strategies to gather more labeled data could significantly improve model performance and generalizability.
 

### 2025-04-15 paper/paper.qmd
Final

--- 2025-04-15 paper/paper.qmd
@@
 
-Strong lensing occurs when a massive foreground object, such as a galaxy cluster, bends light from a more distant source, producing distinctive arcs or rings in images.
-Despite their significance for cosmology, these events are extremely rare, making up only a small fraction of astronomical observations.
-Traditional supervised learning methods require large quantities of labeled data, which is unfeasible in this context.
-The primary challenge, therefore, is to develop a system capable of identifying potential lensing candidates with minimal labeled data.
+Strong gravitational lensing occurs when a massive foreground object, such as a galaxy cluster, bends the light from a more distant source, producing distinctive arcs or rings in astronomical images.
+Although these phenomena are of great importance to cosmology, they are extremely rare—accounting for only a small fraction of all astronomical observations.
+Traditional supervised learning approaches typically require large amounts of labeled data, which is not feasible in this context due to the scarcity of strong lensing examples.
 
-Thus we will use the state of art active learning strategy, BADGE, which aim to selects the most informative data points for human annotation and thus would improve training efficiency of the Machine Learning model [@cite].
-This strategy have already had promising outcomes in many lacking labeled classification tasks and is of the most effective ways in active learning [@cite].
+To overcome this challenge, we adopt a state-of-the-art active learning strategy known as BADGE, which selects the most informative data points for human annotation.
+This method aims to improve training efficiency by prioritizing samples that are both uncertain and diverse [@cite].
+BADGE has shown strong performance in other low-label settings and is recognized as one of the most effective techniques in active learning [@cite].
 
-However, we are still facing another problem which is the imbalanced data.
-the strong lensing ones in our dataset only consists 8% of the total data.
-And thus will make the evaluation and traing very trick.
-So we will incorporate better evaluating metrics and training loss.
+Despite this, we still face a significant issue: class imbalance.
+In our dataset, strong lensing instances represent only about 8% of the total data.
+This imbalance poses difficulties for both model training and evaluation.
+To address this, we integrate loss functions and evaluation metrics specifically designed for imbalanced classification tasks.
 
-The remainder of this paper is organized as follows.
-@sec-method describes possible losses functions we incorporate into our model.
-@sec-metric introduces the new metrics we incorporated @sec-results presents the results of new losses and evaluation metric.
-Finally, @sec-diss discusses the implications of the findings, the limitations of the study, and recommendations for future research in finding strong lensing data.
+The remainder of this paper is structured as follows: Section @sec-method outlines our approach, including the active learning framework, loss functions, and evaluation metrics used.
+Section @sec-results presents the performance results under the proposed method.
+Finally, Section @sec-diss discusses the implications of our findings, current limitations, and directions for future research in detecting strong lensing events.
 
@@
 
-To test the new metric and loss functions, we use the available strong lensing data for regression.
-We concatenate the features extracted from the CNN and contatenate them with the expert labels for regression.
-The result data has a total of ~ and ~ strong lenseing ones.
+To evaluate the effectiveness of the proposed loss functions and evaluation metric, we use available strong lensing data in a regression framework.
+Features are first extracted using a convolutional neural network (CNN), and then concatenated with expert-provided labels for training.
+The resulting dataset contains approximately ~ total samples, with ~ representing strong lensing cases.
 
@@
 
-Here we will investigate two loss functions, ranknet loss and focal loss.
-Ranknet loss is sepcified as follow: Given training data $x_1$, $x_2$, ....
-, $x_n$, and their rankings, $y_1$, $y_2$, ..., $y_n$, we will investigate all combinations of ordered pairs ($x_i$, $x_j$) where their labels $y_i$ and $y_j$ are different.
-The total loss is:
+We explore two specialized loss functions: RankNet loss and Focal loss.
+
+### RankNet Loss:
+RankNet loss is designed to capture the relative ordering between data points based on their labels.
+Given training data points $x_1, x_2, \ldots, x_n$ with corresponding rankings $y_1, y_2, \ldots, y_n$, we consider all ordered pairs $(x_i, x_j)$ where $y_i \ne y_j$.
+The loss is computed as:
 $$
@@
 $$
-Where $y_ij$ is 1 if $x_i$ is ranked higher than $y_j$, 0 if $x_i$ is ranked lower than $y_j$ (here we will assume no $x_i$ can have the same ranking as $x_j$ if i is not equal to j).
-$P(i > j) = \frac{1}{1 + e^{-(s(x_i) - s(x_j))}}$ is the probability that $x_i$ is ranked higher than $x_j$.
-s() is the rankning score assigned to i, higher ranking cores means higher ranking.
-This loss aims to peanalize the pairs of data $x_i$ and $x_j$ where $x_i$ should be ranked higher than $x_j$ but instead $x_i$ is ranked lower by the model for $x_j$.
+Here, $y_{ij} = 1$ if $x_i$ is ranked higher than $x_j$, and $0$ otherwise (assuming no ties in ranking).
+The probability that $x_i$ should be ranked above $x_j$ is defined as: $$ P(i > j) = \frac{1}{1 + e^{-(s(x_i) - s(x_j))}} $$.
+where $s(\cdot)$ is the model’s predicted ranking score.
+This loss penalizes the model when it incorrectly ranks a more relevant (higher-lensed) sample lower than a less relevant one.
 
-The focal loss is specified as follows: 
+### Focal Loss:
+Focal loss is particularly effective in handling class imbalance by focusing learning on hard-to-classify examples. It is defined as:
 $$
@@
 $$
-Here $p_t$ is the model's predicted probility if the true label is strong lensing, one minus the predicted probbility if the true label is not strong lenseing.
-$a_t$ is 0.75 if the true lavel is strong lensing and 0.25 if the true label is not strong lensing.
-The aim of this loss function is to avoid easy labels and penalize the model most if it incorrectly label a strong lensing one as non-lensing.
+Here, $p_t$ is the predicted probability for the true class: the model’s predicted probability if the true label is strong lensing, and $1 - p_t$ otherwise.
+The weighting factor $\alpha_t$ is set to 0.75 for strong lensing samples and 0.25 for non-lensing ones.
+This loss down-weights well-classified examples and puts more emphasis on correcting misclassified strong lensing instances, which are more valuable in our context.
 
@@
 
-Despite the fact that we need new loss functions, we also need new evaluation metrics to better evaluate the performance of model on biased labels.
-As said, the labels of strong is very rare relative to non lensing.
-So if we use the default prediction accuracy then by just predicting everything by strong lensing the model will get a very decent performance while did not contribute much to the final goal of our research purpose.
+Due to the severe class imbalance—strong lensing events account for only ~8% of the dataset—traditional evaluation metrics like overall accuracy can be misleading.
+For example, a model that predicts all samples as non-lensing would achieve high accuracy, yet fail to serve our scientific objectives.
 
-Thus we will propose a practical and better to understand metric, which is the # of correct labels in the top 10% of the highest ranked data.
-This is more to common sense and align the final purpose of our research.
+To better reflect the model's practical performance, we propose a more intuitive and goal-aligned metric:
+the number of correctly identified strong lensing samples within the top 10% of the highest-ranked predictions.
 
-# Results
+This metric directly measures the model's ability to surface valuable candidates for human review and is more consistent with the real-world usage of the system—prioritizing rare but scientifically significant events for expert validation.
 
-To test the two loss functions useing the new metric we build a one layer neuronetwork with 40 input neurons, 128 hidden neurons and one output which is the score of the input data in pytorch.
-We split the dataset into training and testing parts with the ratio of 7:3.
-To gain efficiency in training, we split the training data into minibatches of size 128 and after finetuning set the training epohs be 100 to avoid overtraining.
-We also used a neuronet work with crossentropy loss as a conterpart.
-We ran 10 independent groups and the result is the following:
+# Results {#sec-Results}
+
+To evaluate the two proposed loss functions using the new metric, we built a simple neural network in PyTorch.
+The model architecture consists of an input layer with 40 neurons, one hidden layer with 128 neurons, and a single output neuron representing the predicted score of the input sample.
+The dataset was split into training and testing sets with a 7:3 ratio.
+To improve training efficiency, we used mini-batches of size 128 and set the number of training epochs to 100 to avoid overfitting based on preliminary fine-tuning.
+
+For comparison, we also trained a baseline model using standard cross-entropy loss. All models were trained and evaluated across 10 independent runs using different random seeds.
 
@@
 
-Over the past 30 years, advancements in society and technology have significantly reduced global mortality rates, particularly for children under the age of five (child mortality).
-The child mortality rate declined from 93 deaths per 1,000 live births in 1990 to 37 deaths per 1,000 live births in 2022 [@worldbank2024M].
-This progress represents millions of children who now have better chances of survival, marking a substantial achievement in global health.
+@fig-mor1 presents the average number of correctly identified strong lensing instances among the top 100 ranked predictions (approximately the top 10% of the test set).
+This metric directly reflects the model's ability to surface the most relevant candidates.
 
-Despite this progress, the pace of reduction has slowed considerably since 2015, decreasing from an annual rate of 3.8% during the Millennium Development Goal (MDG) era (2000–2015) to 2.1% during the Sustainable Development Goal (SDG) era (2015–2022) [@unicef2022underfive].
-In 2022 alone, 4.9 million children under the age of five died, equating to approximately 13,400 deaths per day [@unicef2022underfive].
-Moreover, the distribution of mortality rates under five remains highly uneven.
-Developed countries such as Canada and the United States report mortality rates as low as 10 per 1,000 live births while developing nations—particularly in sub-Saharan Africa—experience rates exceeding 100 per 1,000 live births [@worldbank2024M].
-This stark disparity underscores the urgent need for targeted interventions in high-mortality regions.
+We observe that all three loss functions—cross-entropy (MLP), RankNet, and Focal Loss—perform similarly, with the average number of correct predictions hovering around 40.
+While RankNet does not show clear improvement over the baseline, Focal Loss slightly outperforms cross-entropy in some runs.
+However, the differences are not statistically significant, as indicated by the overlapping error bars.
 
-Infectious diseases, such as pneumonia, diarrhea, and malaria, remain the leading causes of under-five deaths, closely tied to vaccination coverage and healthcare spending.
-Maternal health is also influenced by healthcare accessibility and quality, which often correlates with national health expenditures [@who2024childmortality].
-While food production does not directly address these issues, it reflects a nation’s ability to meet nutritional needs, a critical factor in preventing malnutrition—a major contributor to child mortality.
+Overall, while Focal Loss shows some promise, none of the alternative loss functions demonstrated a substantial performance gain under this evaluation setting.
 
-In this study, we examined the child mortality rate for each country in 2022, focusing on three key factors: food production index, DPT vaccine coverage, and per-capita health expenditure.
-Our findings reveal that increasing vaccine coverage and per-capita health expenditures significantly reduce under-five mortality rates.
-However, the food production index does not exhibit a direct impact, suggesting its role is mediated by factors such as food quality and distribution systems.
+# Dsicussion {#sec-discussion}
 
-The estimand of this study is the effect of three predictors—food production index, vaccination coverage (DPT vaccine percentage), and per-capita health expenditure (log-transformed)—on the logarithm of the under-five mortality rate for each country in 2022.
-This analysis seeks to quantify the extent to which these factors contribute to variations in mortality rates across nations.
-By addressing these relationships, the study aims to identify potential intervention points for reducing preventable child deaths and achieving progress toward the Sustainable Development Goals for child survival.
+## Conslusion and limitation
 
-The remainder of this paper is organized as follows.
-@sec-data describes the data sources, key variables, and preprocessing steps, including transformations applied to address skewness and improve model validity.
-@sec-model outlines the modeling approach, including the choice of predictors, justification of the model structure, and Bayesian implementation.
-@sec-results presents the results of the analysis, highlighting the effects of each predictor on under-five mortality.
-Finally, @sec-diss discusses the implications of the findings, the limitations of the study, and recommendations for future research and policy interventions aimed at reducing child mortality globally.
+Based on our experiments, Focal Loss appears to be the most promising option among the three loss functions tested.
+However, in practice, it introduces increased computational complexity during training—particularly when used in combination with active learning strategies like BADGE, which requires computing gradients with respect to the penultimate layer of the neural network.
 
-# Data {#sec-data}
+With Cross-Entropy Loss, the gradient has a simple, closed-form expression, which makes it computationally efficient and easier to integrate into BADGE.
+In contrast, Focal Loss includes additional terms involving $\alpha_t$ and $\gamma$, which complicate gradient calculations and increase training time.
+To make Focal Loss more viable in practice, it may be necessary to fine-tune its hyperparameters to improve both performance and efficiency.
 
-## Overview
+As for RankNet Loss, its application in our context is limited by the rarity of training pairs with differing labels in each mini-batch.
+Since strong lensing cases are so scarce, many batches end up containing only one class, making it impossible to form valid comparison pairs.
+This severely hampers the effectiveness of the RankNet approach under class-imbalanced conditions.
 
-For this analysis, we used four datasets: child mortality rate, food production index, DPT vaccine coverage rate, and health expense per capita.
-All four datasets are downloaded from **Worldbank Open Database** [@worldbank_data] through its API [@worldbank].
-World Bank Open Database is a comprehensive resource offering free and publicly accessible development data from nearly every country.
-This platform provides an extensive range of indicators across sectors such as health, agriculture, and education, enabling researchers to access globally standardized data for comparative analysis.
-The platform ensures transparency and usability by offering tools for visualization, data export, and API integration, making it a reliable and versatile resource for exploring global trends.
+## Future work
 
-We employed **R** [@citeR], a coding platform to download, clean, and conduct statistical analysis.
-Besides, we also utilized R packages **tidyverse** [@tidy], **rstanarm** [@rstanarm], **ggplot2** [@gg], **knitr** [@knitr], **arrow** [@arrow], **here** [@here], **modelsummary** [@ms], **testthat** [@testthat] and **dpylr** [@dplyr].
-The paper is outlined in Git Hub using the starter folder provided in **Telling Stories With Data** [@tellingstories].
+While our current study highlights some key findings, it also reveals several opportunities for further research and development.
+One promising direction is to fine-tune the CNN feature extractor used to generate input representations for the model.
+In our experiments, the CNN was kept fixed, but optimizing it alongside the main model could lead to more expressive features and improved performance in identifying strong lensing events.
 
-## Measurement
+Another avenue is the exploration of alternative loss functions or ranking strategies that are more suitable for highly imbalanced datasets.
+For example, newer ranking-based methods or hybrid loss formulations might offer better trade-offs between learning signal and computational complexity, especially when valid data pairs are scarce—as was the case with RankNet in our experiments.
 
-The datasets of this analysis are collected and prepared by various UN-affiliated organizations.
-Below, we provide a brief overview of the measurement and transformation processes for these data sets.
-Detailed descriptions and methodological specifics can be found in @sec-methods.
+In addition, refining the evaluation metric to better reflect real-world objectives would strengthen the connection between model outputs and practical use.
+While our current metric—correct predictions in the top 10%—is more aligned with human review processes, future work could explore even more nuanced measures of model utility in scientific workflows.
 
-### Response variable:
+Finally, incorporating advanced active learning strategies represents a critical path forward.
+By intelligently selecting which samples to label, we can increase the diversity and informativeness of the training data while minimizing labeling effort.
+This is especially relevant given the scarcity of strong lensing cases and the high cost of expert annotation.
 
-**Mortality Rate**: The child mortality rate is estimated by the UN Inter-agency Group for Child Mortality Estimation (IGME).
-The primary data source is each country's vital registration system.
-In countries where such a system is lacking, household surveys are conducted and then processed using a statistical model developed collaboratively by UN agencies and academic institutions to derive an estimate of the mortality rate [@worldbank2024M].
-
-### Predictors:
-
-**Food Production Index**: This index, prepared by the Food and Agriculture Organization (FAO) of the United Nations, measures aggregate food production relative to the 2014–2016 baseline period.
-It is calculated yearly using price-weighted data on agricultural products for each country.
-[@worldbank_foodindex2024].
-
-**Health Expenditure per Capita**: Data on health expenditure per capita are produced by the World Health Organization (WHO) following the System of Health Accounts 2011 (SHA 2011) framework.
-This framework tracks all health-related spending in a country over a defined period [@worldbank_health_expenditure_us].
-
-**DPT Vaccine Coverage**: DPT vaccine coverage is jointly estimated by the WHO and UNICEF.
-This metric reflects the percentage of children aged 12–23 months in each country who have received diphtheria, pertussis (whooping cough), and tetanus (DPT) vaccine.
-Estimates are derived from reports by vaccine service providers and household surveys of vaccination history [@worldbank_dpt_immunization].
-
-## Outcome variables
-
-The response variable in this study is the child mortality rate in each country in 2022, measured as the number of deaths of children under five years old per 1,000 live births in 2022.
-
-The Original data range from 0 to 120, with a peak between 10 and 15 deaths per 1,000 live births (@fig-mor-1).
-Due to extreme values and significant right skewness, a log transformation was applied to stabilize variance and improve normality.
-The transformed data range from 2 to 5, with a peak near 3, and the shape is approximately normal.
-(@fig-mor-2).
-
-```{r}
-#| label: fig-mor
-#| fig-cap: Data Analysis for Mortality Rate for children under five in different countries.
-The histogram on the top left shows that the original distribution is skewed to the right.
-The histogram on the top right is the distribution after log transformation and is approximately normal.
-#| fig-subcap: ["Distribution of Mortality rate before log transformation", "Distribution of Mortality rate after log transformation"]
-#| echo: false
-#| warning: false
-#| message: false
-#| layout-ncol: 2
-
-# Read the data
-data <- read.csv(file = (here::here("data/01-raw_data/raw_data_Mortality.csv")))
-
-# Create the histogram
-ggplot(data, aes(x = SH.DYN.MORT)) +
-  geom_histogram(binwidth = 7, fill = "skyblue", color = "black", alpha = 0.7) +
-  labs(
-    x = "Under-Five Mortality Rate",
-    y = "Frequency"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-    axis.title = element_text(size = 12)
-  )
-
-analysis_data <- read_parquet(file = (here::here("data/02-analysis_data/analysis_data.parquet")))
-
-# Create the histogram
-ggplot(analysis_data, aes(x = Mortality)) +
-  geom_histogram(binwidth = 0.3, fill = "skyblue", color = "black", alpha = 0.7) +
-  labs(
-    x = "Under-Five Mortality Rate",
-    y = "Frequency"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-    axis.title = element_text(size = 12)
-  )
-```
-
-## Predictor variables
-
-The three predictors analyzed are:
-
-**Food Production Index:** 
-The Food Production Index measures the relative level of agricultural production for each country in 2022, using the 2014–2016 baseline as a reference point [@worldbank].
-This index is price-weighted and reflects agricultural productivity across countries.
-As illustrated in @fig-food-1, the index ranges from 60 to 180, with a mean of around 110 and an approximately normal distribution.
-The scatter plot between the food production index and log of child mortality in @fig-food-2 shows a slightly positive linear relationship.
-This relationship is examined in greater detail in @sec-results.
-```{r}
-#| label: fig-food
-#| fig-cap: Data Analysis of Predictor Food Production Index.
-The histogram on the left shows an approximately normal distribution.
-The scatterplot on the right indicates there is a slight positive relationship between food production index and mortality rate.
-#| fig-subcap: ["Counts of Food production index for each country, measured as a basis of 2014-1016", "Relationship of each countries' Food production index VS. Log of Mortality rate"]
-#| echo: false
-#| warning: false
-#| message: false
-#| layout-ncol: 2
-
-analysis_data <- read_parquet(file = (here::here("data/02-analysis_data/analysis_data.parquet")))
-
-# Create the histogram
-ggplot(analysis_data, aes(x = Food)) +
-  geom_histogram(binwidth = 5, fill = "skyblue", color = "black", alpha = 0.7) +
-  labs(
-    x = "Food Production Index",
-    y = "Frequency"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-    axis.title = element_text(size = 12)
-  )
-
-ggplot(analysis_data, aes(x = Food, y = Mortality)) +
-  geom_point(color = "steelblue", alpha = 0.6) +
-  geom_smooth(method = "lm", color = "darkred", fill = "pink", se = TRUE, alpha = 0.3) +
-  labs(
-    x = "Food Production Index",
-    y = "Mortality Rate",
-  ) +
-  theme_minimal() +
-  theme(
-    axis.title = element_text(size = 12)
-  )
-```
-**Current Health Expenditure:**
-The Current Health Expenditure per Capita captures health spending in US dollars for each individual in 2021, calculated under the System of Health Accounts (SHA 2011) framework.
-The raw data range from $0 to $12,000 and exhibit extreme right skewness, as shown in @fig-health-1.
-To address this, a log transformation was applied, resulting in a normalized range from 0 to 10, with a central tendency near 6 (@fig-health-2).
-Analysis of the relationship between log-transformed health expenditure and log mortality rate (@fig-health-3) demonstrates a significant negative linear trend, indicating that higher health expenditures are associated with lower mortality rates.
-This relationship is examined in greater detail in @sec-results.
-```{r}
-#| label: fig-health
-#| fig-cap: Data Analysis of health expenses per capita.
-The histogram on the top left shows that the original distribution is skewed to the right.
-The histogram on the top right is the distribution after log transformation and is approximately normal.The scatterplot on the bottom left indicates there is a negative relationship between log of health expenses and Mortality rate.
-#| fig-subcap: ["Counts of Health Expenditure Per Capita for each country (measured in US$) before log transformation", "Counts of each country's Health Expenditure Per Capita, (measured in US$) after log transformation", "Relationship of each countries' Health Expenditure Per Capita, Measured in US$ VS.
-Log of Mortality"]
-#| echo: false
-#| warning: false
-#| message: false
-#| layout-ncol: 2
-
-raw_data <- read.csv(file = (here::here("data/01-raw_data/raw_data_Healthexpense.csv")))
-
-analysis_data <- read_parquet(file = (here::here("data/02-analysis_data/analysis_data.parquet")))
-
-ggplot(raw_data, aes(x = SH.XPD.CHEX.PC.CD)) +
-  geom_histogram(binwidth = 500, fill = "skyblue", color = "black", alpha = 0.7) +
-  labs(
-    x = "Health Expense per capita",
-    y = "Frequency"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-    axis.title = element_text(size = 12)
-  )
-
-# Create the histogram
-ggplot(analysis_data, aes(x = Health_expense)) +
-  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black", alpha = 0.7) +
-  labs(
-    x = "Log of Health Expense per capita",
-    y = "Frequency"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-    axis.title = element_text(size = 12)
-  )
-
-ggplot(analysis_data, aes(x = Health_expense, y = Mortality)) +
-  geom_point(color = "steelblue", alpha = 0.6) +
-  geom_smooth(method = "lm", color = "darkred", fill = "pink", se = TRUE, alpha = 0.3) +
-  labs(
-    x = "Log of Health Expense per capita",
-    y = "Mortality Rate",
-  ) +
-  theme_minimal() +
-  theme(
-    axis.title = element_text(size = 12)
-  )
-```
-**DPT Vaccine Coverage:**
-DPT vaccine coverage rate reflects the percentage of children aged 12–23 months who received vaccinations against diphtheria, pertussis, and tetanus in 2022.
-Coverage ranges from 30% to 100%, with a left-skewed distribution, as depicted in @fig-DPT-1.
-Analysis of the association between vaccine coverage and mortality rates (@fig-DPT-2) shows a strong negative relationship, suggesting that increased vaccine coverage is correlated with lower mortality rates.
-This finding will be explored further in @sec-model.
-```{r}
-#| label: fig-DPT
-#| fig-cap: Data Analysis of DPT Vaccine Coverage.
-The histogram on the left shows that the distribution of DPT vaccine coverage is skewed to the left.
-The scatterplot on the right indicates there is a negative relationship between DPT vaccine coverage and Mortality rate.
-#| fig-subcap: ["Counts of each country's DPT vaccine percentage", "Relationship of each country's DPT Vaccine Percentage VS. Log of Mortality"]
-#| echo: false
-#| warning: false
-#| message: false
-#| layout-ncol: 2
-
-analysis_data <- read_parquet(file = (here::here("data/02-analysis_data/analysis_data.parquet")))
-
-# Create the histogram
-ggplot(analysis_data, aes(x = Vacinne)) +
-  geom_histogram(binwidth = 5, fill = "skyblue", color = "black", alpha = 0.7) +
-  labs(
-    x = "DPT vaccine percentage",
-    y = "Frequency"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-    axis.title = element_text(size = 12)
-  )
-
-ggplot(analysis_data, aes(x = Vacinne, y = Mortality)) +
-  geom_point(color = "steelblue", alpha = 0.6) +
-  geom_smooth(method = "lm", color = "darkred", fill = "pink", se = TRUE, alpha = 0.3) +
-  labs(
-    x = "DPT vaccine percentage",
-    y = "Mortality Rate",
-  ) +
-  theme_minimal() +
-  theme(
-    axis.title = element_text(size = 12)
-  )
-```
-## Missing Data and Time Inconsistency {#sec-data-missing}
-The health expenditure variable was collected in 2021, while other variables reflect 2022 data.
-This discrepancy is due to the unavailability of more recent health expenditure data.
-These 2021 values are assumed to approximate 2022 conditions sufficiently for analysis.
-
-Additionally, some countries' data were not collected and were excluded from the analysis. The implications of these omissions and time inconsistency are addressed in @sec-lim.
-
-# Model {#sec-model}
-
-Our modeling strategy aims to explore the relationships between DPT vaccine coverage, the food production index, and current health expenditure per capita (in current US$) with the under-five mortality rate across nations.
-By analyzing these factors, we seek to identify opportunities to improve child health outcomes, particularly in countries with high mortality rates.
-Supporting model diagnostics are provided in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Let $y_i$ be the logarithm of the under-five mortality rate for nation $i$. We define the following predictors:
-
-$x_{1i}$: DDT vaccine coverage for nation $i$.
-$x_{2i}$: Food production index for nation $i$.
-$x_{3i}$: Logarithm of current health expenditure per capita (current US$) for nation $i$.
-The model is specified as:
-
-\begin{align} 
-y_i \mid \mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-u_i &= \alpha + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i} \\
-\alpha &\sim \mbox{Normal}(2.5, 1) \\
-\beta_1 &\sim \mbox{Normal}(0, 2.5)\\
-\beta_2 &\sim \mbox{Normal}(0, 2.5)\\
-\beta_3 &\sim \mbox{Normal}(0, 2.5)\\
-\sigma &\sim \mbox{Exponential}(1) 
-\end{align}
-
-We implement this model in **R** [@citeR] using the **rstanarm**  [@rstanarm] and **arrow** [@arrow] package.
-
-### Model justification
-
-A Bayesian multivariate linear model was chosen to capture the relationships between the predictors (DPT vaccine coverage, food production index, and current health expenditure) and the response variable (under-five mortality rate).
-This choice is supported by observed linear trends between each predictor and response in @sec-data.
-Logarithmic transformations were applied to the under-five mortality rate and current health expenditure per capita to normalize the distributions and stabilize their linear relationships with the predictors.
-
-Default Bayesian priors were used for the predictors, as they provide a robust starting point.
-For the intercept term, a \mbox{Normal}(2.5, 1) prior was chosen, reflecting the observed distribution of the response variable (@fig-mor-1).
-
-### Model weakness
-Several limitations are noted in this model:
-
-1. Skewed Predictor Distribution: The distribution of DPT vaccine coverage is skewed left, which may introduce bias. Despite attempts to address this, an ideal solution remains elusive.
-
-2. Potential Collinearity: While each predictor shows an individual linear relationship with the response variable, dependencies between predictors may affect model accuracy.
-
-3. Simplified Assumptions: The model assumes linearity and does not capture potential non-linear interactions or regional variations that could influence under-five mortality rates.
-
-## Model Validation
-The posterior and prior predictive checks are presented in @fig-ppcheckandposteriorvsprior-1.
-We observe an excellent fit between the posterior distribution from our model and the actual child mortality data, indicating that the model effectively captures the underlying trends in child mortality rates.
-In @fig-ppcheckandposteriorvsprior-2, we compare the posterior distribution with the prior.
-No significant deviations are observed, suggesting alignment between our prior beliefs and the posterior distribution.
-
-@fig-stanareyouokay-1 displays the trace plot check, where no irregularities or concerning patterns are detected.
-Finally, the R-hat plot (@fig-stanareyouokay-2) shows all values close to one, confirming the convergence of the Markov chains.
-Detailed information could be found in @sec-model-details.
-
-# Results {#sec-results}
-
-## Model outcome
-
-Our model results are summarized in @tbl-modelresults. We will discuss them one by one.
-
-### Intercept:
-The intercept estimate of 6.643 represents the expected logarithm of the under-five mortality rate when all predictors are held constant at their reference or baseline values (e.g., average food production, average vaccine coverage, and average health expenditure).
-
-### Food Production Index:
-The coefficient for the food production index is 0.003, indicating a very slight positive relationship between food production and the under-five mortality rate.
-This naively suggests that reducing food production could also reduce mortality, which is counter-intuitive.
-
-However, a closer examination of the confidence interval for this predictor shows that it includes 0.
-This implies that the relationship is not statistically significant, so we conclude that food production is not strongly related to overall under-five mortality rates.
-
-### Vaccine Coverage:
-The coefficient for vaccine coverage is -0.013, meaning that for every 1 percentage point increase in vaccine coverage, the logarithm of the under-five mortality rate decreases by 1.3%, holding other variables constant.
-
-### Health Expenditure (Log-Transformed):
-The coefficient for the log-transformed health expenditure is -0.529, indicating that a 1% increase in health expenditure per capita is associated with a 0.529 unit decrease in the log of the under-five mortality rate, holding all other predictors constant.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory model of Mortality rate for each country based on food production index, DPT vaccine coverage, and health expense."
-
-#### Load the saved model ####
-# Load the saved models
-model1 <- readRDS(file = here::here("models/Mortality_Model.rds"))
-
-modelsummary(list("Mortality rate model" = model1), fmt = 4, statistic = c("({conf.low}, {conf.high})"))
-```
-
-# Discussion {#sec-diss}
-
-## Overview
-The United Nations’ Sustainable Development Goal (SDG) 3.2 aims to reduce the mortality rate of children under five to fewer than 25 per 1,000 live births by 2030 [@who2024].
-While substantial progress has been achieved globally, meeting this ambitious target within the remaining six years presents significant challenges.
-Although global trends indicate a decline in child mortality rates, the pace of this reduction has slowed since 2015 [@who2024].
-
-A stark disparity persists across countries, as illustrated in @fig-mor-1.
-While some nations, especially high-income ones, have already surpassed the SDG target, others—predominantly in low-income regions such as sub-Saharan Africa—experience child mortality rates that are four times higher than the global goal.
-
-To achieve the SDG targets, concerted efforts are essential, particularly through targeted support from international organizations like the World Health Organization (WHO).
-However, this raises critical questions: Where should resources be allocated most effectively?
-How can countries with high child mortality rates achieve the greatest impact with limited resources?
-
-## Correlation V.S. Causality
-
-Numerous studies have explored the determinants of child mortality, consistently finding a negative correlation between mortality rates and macroeconomic indicators such as GDP [@ohare2013income] or education levels [@sartorius2014global].
-However, many of these factors are not direct causes of child mortality.
-
-For instance, during the Great Recession, the United States experienced a reduction in mortality rates due to decreased air pollution following widespread factory closures [@finkelstein2024lives].
-This demonstrates the importance of identifying underlying causes rather than focusing solely on broad economic indicators.
-In low-income countries, the primary drivers of high child mortality rates are often basic issues such as malnutrition and inadequate access to primary healthcare [@DOERR2022110860].
-Addressing such root causes is crucial to achieving meaningful and sustainable improvements.
-
-Our analysis identifies two factors with strong negative correlations to child mortality rates: DPT vaccine coverage and healthcare spending.
-However, it remains essential to determine whether these factors exert a direct influence or are proxies for other underlying variables.
-
-The DPT vaccine protects against three deadly diseases: diphtheria, pertussis (whooping cough), and tetanus (lockjaw).
-Administered in five doses starting at two months of age, this vaccine provides critical immunity for newborns [@cdc_dtap].
-
-In 1990, diphtheria, pertussis, and tetanus collectively caused approximately 160,000 deaths among children aged 1–4 years.
-By 2019, this number had decreased to 60,000, demonstrating the vaccine's impact [@gbd2019].
-Nonetheless, disparities in vaccination rates persist, leaving many children vulnerable to preventable diseases.
-Studies indicate that children receiving zero doses of vaccines are at significantly higher risk of mortality [@karlsson2024prevalence].
-Expanding vaccination coverage is a direct and cost-effective intervention to reduce child mortality and improve public health.
-
-Healthcare spending encompasses all expenditures on health-related goods and services within a given year, including both private and public spending [@worldbank_health_expenditure_us].
-Research suggests that public healthcare spending has the most pronounced effect on reducing child mortality [@ray2020health], primarily by increasing access to trained healthcare professionals and improving maternity and neonatal care.
-
-Public healthcare systems are often funded through tax revenues, which are tied to broader economic performance and governance structures [@ray2020health].
-While boosting overall tax revenues may be challenging, reallocating existing resources to prioritize public health can yield significant benefits.
-This approach is particularly relevant in countries where public spending on healthcare remains disproportionately low.
-
-## Conclusion
-
-Our study shows that reducing the child mortality rate requires a dual focus on expanding vaccination coverage and optimizing healthcare spending.
-Vaccination programs, particularly for diseases like diphtheria, pertussis, and tetanus, represent a direct strategy for reducing child mortality.
-Simultaneously, enhancing public healthcare spending can address systemic issues such as inadequate maternal care and limited access to trained professionals.
-
-On the other hand, efforts must be tailored to the specific needs and capacities of individual countries.
-International organizations can play a pivotal role by providing financial support, technical expertise, and policy guidance to countries with high child mortality rates.
-By prioritizing interventions with clear and measurable impacts, the global community can move closer to the goal of ensuring that no child dies from preventable causes.
-
-## Limitation and next steps {#sec-lim}
-This study has several limitations that warrant consideration in future research.
-First, as discussed in @sec-data-missing, data from several countries is unavailable.
-While these nations typically have smaller populations, their exclusion may introduce biases and raise ethical concerns by neglecting their unique challenges.
-Second, the most recent data used in this study is from 2022, and there are temporal inconsistencies among the predictors.
-This is particularly important because both the predictors and the response variable are time-sensitive.
-Access to more current and synchronized datasets would enable a more accurate and relevant analysis, reflecting recent developments in healthcare and child mortality trends.
-
-Additionally, while this study focused on three primary predictors, the dataset contains other indicators that could potentially enhance the analysis.
-However, many of these indicators are affected by missing or outdated values, limiting their utility.
-Future research should prioritize addressing data completeness and timeliness to improve model robustness and reliability.
-
-Finally, incorporating advanced modeling techniques and additional predictors could provide deeper insights.
-Future studies should explore innovative models to better capture complex relationships and interactions, enabling more nuanced and actionable recommendations for policy interventions.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Sampling Methodology Overview and Evaluation {#sec-methods}
-
-## Overview
-
-This study explores the relationship between under-five mortality rates and three key predictors: food production index, DPT vaccine coverage, and per capita health expenditure.
-The under-five mortality rate is derived from a combination of vital registration systems, household surveys (e.g., Demographic and Health Surveys and Multiple Indicator Cluster Surveys), and indirect estimation models developed by UN agencies and collaborating academic institutions [@worldbank2024M].
-These data sources aim to address disparities in coverage, particularly in regions where vital registration is incomplete.
-
-The food production index, developed by the Food and Agriculture Organization of the United Nations (FAO), quantifies national agricultural output relative to a baseline period (2014–2016) using official national statistics and supplementary surveys [@worldbank_foodindex2024].
-Per capita health expenditure is sourced from the World Health Organization (WHO) using the System of Health Accounts (SHA 2011), which systematically tracks all public and private spending on health services, goods, and infrastructure [@worldbank_health_expenditure_us].
-Lastly, DPT vaccine coverage is estimated through a combination of administrative data from immunization service providers and household surveys, reflecting the proportion of children aged 12–23 months who received three doses of the vaccine [@worldbank_dpt_immunization].
-
-## Population sample frame, and sampling approach
-
-The target population for the response variable, under-five mortality rate, includes all children under the age of five globally in 2022.
-This metric captures the probability per 1,000 live births that a newborn will die before reaching five years of age, providing a direct measure of child survival across regions and contexts.
-The predictors each represent distinct populations.
-The food production index measures agricultural output for entire national populations, reflecting food availability and potential nutritional impacts.
-Similarly, per capita health expenditure pertains to the entire population within a country, as healthcare spending influences system-wide access and quality.
-DPT vaccine coverage, by contrast, specifically targets the population of children aged 12–23 months, providing insights into immunization rates and public health outreach.
-
-The sampling frame for each variable aligns with its respective population.
-Under-five mortality rates rely on vital registration systems, which are designed to record all births and deaths.
-In regions where these systems are incomplete, surveys and statistical models supplement the data to ensure national representativeness.
-The food production index uses national agricultural statistics as the primary source, with secondary data from government ministries and international organizations filling gaps [@worldbank_foodindex2024].
-For DPT vaccine coverage, data are collected through administrative immunization records verified by household surveys [@worldbank_dpt_immunization].
-Per capita health expenditure is derived from national accounts and follows the standardized SHA 2011 framework [@worldbank_health_expenditure_us].
-
-The sampling approach for these variables combines census-based data collection with stratified sampling in regions lacking comprehensive infrastructure.
-While census methods aim to capture complete population-level data, resource constraints in low- and middle-income countries necessitate reliance on surveys and indirect estimation methods.
-These approaches use stratification to ensure representativeness while accounting for regional disparities.
-
-## Strength and weakness
-
-The methodological framework demonstrates several strengths.
-First, the primary reliance on census-based or near-census-level data ensures comprehensive coverage for most variables.
-This is particularly evident in high-income regions, where systems such as vital registration achieve over 90% completeness [@WHO2024CRVS].
-Second, the use of supplementary data sources, such as surveys or administrative records, mitigates gaps in the primary datasets, enhancing data quality.
-For instance, the food production index excludes non-nutritive products such as coffee and tea, ensuring a focus on agricultural outputs relevant to nutrition [@worldbank_foodindex2024].
-The use of standardized frameworks, such as SHA 2011 for health expenditure, further supports consistency and comparability across countries.
-
-However, notable weaknesses persist.
-In regions with incomplete infrastructure, such as Sub-Saharan Africa, vital registration systems capture only 44% of deaths, leading to substantial data gaps that require imputation through statistical models [@WHO2024CRVS].
-These models rely on assumptions that may not fully reflect local demographic trends, introducing potential biases.
-Moreover, reliance on household surveys for DPT vaccine coverage and mortality data introduces recall bias, as respondents may inaccurately report past events [@Preston1984AgePattern].
-Even census-based data are susceptible to misreporting, undercounting, or misclassification, which can distort results.
-Furthermore, temporal mismatches among variables (e.g., health expenditure data lagging behind mortality rates) may complicate causal inference and trend analysis.
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-We conducted a posterior predictive check to evaluate how well the model predicts under-five mortality rates.
-@fig-ppcheckandposteriorvsprior-1 illustrates the predictive distribution compared to the observed data.
-The close alignment between observed and predicted values indicates that the model captures the main trends in the data effectively.
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
-This shows that the data significantly updates the prior beliefs for all key parameters.
-The posterior distributions of predictors—vaccine coverage, health expenditure, and food production index—highlight the robustness of their effects in the model.
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 1
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-analysis_data <- read_parquet(file = here::here("data/02-analysis_data/analysis_data.parquet"))
-
-pp_check(model1) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(model1) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot.
-It shows adequate mixing of the Markov Chain Monte Carlo (MCMC) chains, indicating proper convergence.
-Additionally, the Rhat values for all parameters (Figure 6b) are below 1.05, further confirming convergence and the reliability of the parameter estimates.
-
-@fig-stanareyouokay-2 is a Rhat plot.
-It shows no significant patterns, indicating that the model assumptions hold.
-Additionally, the Bayesian framework effectively quantifies uncertainty, with credible intervals for the predictors providing insight into their respective influences on mortality.
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 1
-#| fig-width: 8
-
-plot(model1, "trace")
-
-plot(model1, "rhat")
-```
+Together, these directions could significantly enhance the robustness, accuracy, and real-world applicability of models developed for detecting strong gravitational lensing events.
 

### 2025-04-15 scripts/00-simulate_data.R
Restructure

@@ -1,39 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates the data that used to predict the mortality rate in 2022,
-# the predictor variables are food production index, health expenses, and DPT vaccine rate
-# Author: Junbo Li
-# Date: 12 NOV 2024
-# Contact: junb.li@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(arrow)
-
-# Set seed for reproducibility
-set.seed(123)
-
-# Generate data
-n <- 100 # Number of observations
-
-# Generate predictors
-HealthExpenses <- runif(n, 0, 4000) # Health expenses
-DPTVaccineRate <- runif(n, 0, 100) # DPT vaccine rate
-FoodProductionIndex <- runif(n, 10, 100) # Food production index
-
-# Create MortalityRate negatively related to the other variables
-MortalityRate <- 100 - 0.01 * HealthExpenses - 0.3 * DPTVaccineRate - 0.2 * FoodProductionIndex + rnorm(n, mean = 0, sd = 5)
-
-# Combine into a data frame
-dataset <- data.frame(
-  MortalityRate = MortalityRate,
-  HealthExpenses = HealthExpenses,
-  DPTVaccineRate = DPTVaccineRate,
-  FoodProductionIndex = FoodProductionIndex
-)
-
-# Save the dataset as a Parquet file
-write_parquet(dataset, "data/00-simulated_data/simulated_dataset.parquet")

### 2025-04-15 scripts/01-test_simulated_data.R

@@ -1,60 +1,0 @@
-#### Preamble ####
-# Purpose: Tests Structure and validity of simulated data to predicted mortality rate
-# Author: Junbo Li
-# Date: 12 NOV 2024
-# Contact: junb.li@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(arrow)
-library(testthat)
-
-analysis_data <- read_parquet("data/00-simulated_data/simulated_dataset.parquet")
-
-#### Test cases ####
-test_that("Dataset structure and value checks", {
-  # Test if the dataset has 100 rows
-  expect_equal(nrow(dataset), 100, info = "The dataset should have 100 rows.")
-
-  # Test if the dataset has 4 columns
-  expect_equal(ncol(dataset), 4, info = "The dataset should have 4 columns.")
-
-  # Test if 'MortalityRate' values are within the expected range
-  expect_true(all(dataset$MortalityRate >= 5 & dataset$MortalityRate <= 100),
-    info = "'MortalityRate' values should be within the range 5 to 100."
-  )
-
-  # Test if 'HealthExpenses' values are within the expected range
-  expect_true(all(dataset$HealthExpenses >= 0 & dataset$HealthExpenses <= 4000),
-    info = "'HealthExpenses' values should be within the range 0 to 4000."
-  )
-
-  # Test if 'DPTVaccineRate' values are within the expected range
-  expect_true(all(dataset$DPTVaccineRate >= 0 & dataset$DPTVaccineRate <= 100),
-    info = "'DPTVaccineRate' values should be within the range 0 to 100."
-  )
-
-  # Test if 'FoodProductionIndex' values are within the expected range
-  expect_true(all(dataset$FoodProductionIndex >= 10 & dataset$FoodProductionIndex <= 100),
-    info = "'FoodProductionIndex' values should be within the range 10 to 100."
-  )
-
-  # Test if there are no missing values
-  expect_true(all(!is.na(dataset)), info = "The dataset should not contain any missing values.")
-
-  # Test if 'MortalityRate' has at least two unique values
-  expect_true(length(unique(dataset$MortalityRate)) >= 2,
-    info = "'MortalityRate' should have at least two unique values."
-  )
-
-  # Test if there are no negative values in any column
-  expect_true(
[…]

### 2025-04-15 scripts/02-download_data.R

@@ -1,24 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from World Bank opendata
-# Author: Junbo Li
-# Date: 12 NOV 2024
-# Contact: junb.li@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(WDI)
-library(tidyverse)
-#### Download data ####
-data_Mortality <- WDI(indicator = "SH.DYN.MORT", start = 2022, end = 2022)
-data_Food <- WDI(indicator = "AG.PRD.FOOD.XD", start = 2022, end = 2022)
-data_DPT <- WDI(indicator = "SH.IMM.IDPT", start = 2022, end = 2022)
-data_Healthexpense <- WDI(indicator = "SH.XPD.CHEX.PC.CD", start = 2021, end = 2021)
-
-#### Save data ####
-write_csv(data_Food, "data/01-raw_data/raw_data_Food.csv")
-write_csv(data_Mortality, "data/01-raw_data/raw_data_Mortality.csv")
-write_csv(data_DPT, "data/01-raw_data/raw_data_DPT.csv")
-write_csv(data_Healthexpense, "data/01-raw_data/raw_data_Healthexpense.csv")

### 2025-04-15 scripts/03-clean_data.R

@@ -1,41 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the data that is needed for model section
-# Author: Junbo Li
-# Date: 12 NOV 2024
-# Contact: junb.li@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Raw data in data folder
-# Any other information needed? None
-
-#### Workspace setup ####
-# Load the necessary library
-library(dplyr)
-library(tidyverse)
-library(arrow)
-# List of file paths
-file_paths <- c("file1.csv", "file2.csv", "file3.csv", "file4.csv")
-
-# Read and select columns from each file, renaming the unique column
-file1 <- read.csv("data/01-raw_data/raw_data_Food.csv") %>%
-  select(country, Food = AG.PRD.FOOD.XD)
-
-file4 <- read.csv("data/01-raw_data/raw_data_Mortality.csv") %>%
-  select(country, Mortality = SH.DYN.MORT)
-
-file3 <- read.csv("data/01-raw_data/raw_data_DPT.csv") %>%
-  select(country, Vacinne = SH.IMM.IDPT)
-
-file2 <- read.csv("data/01-raw_data/raw_data_Healthexpense.csv") %>%
-  select(country, Health_expense = SH.XPD.CHEX.PC.CD)
-
-# Merge all data frames by 'common_column'
-merged_data <- file1 %>%
-  left_join(file2, by = "country") %>%
-  left_join(file3, by = "country") %>%
-  left_join(file4, by = "country") %>%
-  drop_na() %>%
-  mutate(Health_expense = log(Health_expense)) %>%
-  mutate(Mortality = log(Mortality))
-
-# Save the final merged file
-write_parquet(merged_data, "data/02-analysis_data/analysis_data.parquet")

### 2025-04-15 scripts/04-test_analysis_data.R

@@ -1,60 +1,0 @@
-#### Preamble ####
-# Purpose: Tests Structure and validity of cleaned analysis data to predicted mortality rate
-# Author: Junbo Li
-# Date: 12 NOV 2024
-# Contact: junb.li@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(arrow)
-library(testthat)
-
-dataset <- read_parquet(file = here::here("data/02-analysis_data/analysis_data.parquet"))
-
-#### Test cases ####
-test_that("Dataset structure and value checks", {
-  # Test if the dataset has 100 rows
-  expect_equal(nrow(dataset), 182, info = "The dataset should have 100 rows.")
-
-  # Test if the dataset has 4 columns
-  expect_equal(ncol(dataset), 5, info = "The dataset should have 4 columns.")
-
-  # Test if 'MortalityRate' values are within the expected range
-  expect_true(all(dataset$Mortality >= 0 & dataset$Mortality <= 100),
-    info = "'MortalityRate' values should be within the range 0 to 100."
-  )
-
-  # Test if 'HealthExpenses' values are within the expected range
-  expect_true(all(dataset$Health_expense >= 0),
-    info = "'HealthExpenses' values should be larger than 0."
-  )
-
-  # Test if 'DPTVaccineRate' values are within the expected range
-  expect_true(all(dataset$Vacinne >= 0 & dataset$Vacinne <= 100),
-    info = "'DPTVaccineRate' values should be within the range 0 to 100."
-  )
-
-  # Test if 'FoodProductionIndex' values are within the expected range
-  expect_true(all(dataset$Food >= 0 & dataset$Food <= 500),
-    info = "'FoodProductionIndex' values should be within the range 0 to 500."
-  )
-
-  # Test if there are no missing values
-  expect_true(all(!is.na(dataset)), info = "The dataset should not contain any missing values.")
-
-  # Test if 'MortalityRate' has at least two unique values
-  expect_true(length(unique(dataset$Mortality)) >= 2,
-    info = "'MortalityRate' should have at least two unique values."
-  )
-
-  # Test if there are no negative values in any column
-  expect_true(
-    all(dataset$Health_expense >= 0 & dataset$Vacinne >= 0 &
-      dataset$Food >= 0 & dataset$Mortality >= 0),
-    info = "There should be no negative values in the dataset."
-  )
-})

### 2025-04-15 scripts/05-model_data.R

@@ -1,34 +1,0 @@
-#### Preamble ####
-# Purpose: create and store the Bayesian multi-linear model for this paper
-# Author: Junbo Li
-# Date: 12 NOV 2024
-# Contact: junb.li@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Cleaned data in data folder
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-library(arrow)
-
-#### Read data ####
-analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
-
-# Fit the model using stan_glm with custom priors for scaled predictors
-first_model <- stan_glm(
-  formula = Mortality ~ Food + Vacinne + Health_expense,
-  data = analysis_data,
-  family = gaussian(),
-  prior = normal(location = c(0, 0, 0), scale = c(2.5, 2.5, 2.5), autoscale = FALSE),
-  prior_intercept = normal(location = 2.5, scale = 1, autoscale = TRUE),
-  prior_aux = exponential(rate = 1, autoscale = TRUE),
-  seed = 853
-)
-
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/Mortality_Model.rds"
-)

### 2025-04-15 scripts/06-model_evaluation.R

@@ -1,41 +1,0 @@
-#### Preamble ####
-# Purpose: Obtain an estimate of test score of our model using LOOCV
-# Author: Junbo Li
-# Date: 12 NOV 2024
-# Contact: junb.li@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Cleaned data and model.
-# Any other information needed? None
-
-library(arrow)
-
-# Load dataset
-data <- read_parquet(file = here::here("data/02-analysis_data/analysis_data.parquet"))
-dataset <- data[, 2:4]
-target <- data[, 5]
-
-n <- nrow(dataset)
-
-predictions <- numeric(n)
-actuals <- target
-
-
-model <- read_rds(file = here::here("models/Mortality_Model.rds"))
-
-for (i in 1:n) {
-  # Split data into training and testing sets
-  train_data <- dataset[-i, ]
-  train_target <- target[-i, ]
-  test_data <- dataset[i, , drop = FALSE]
-
-  # Predict for the left-out observation
-  predictions[i] <- predict(model, newdata = test_data)
-}
-
-# Compute LOOCV R-squared
-ss_total <- sum((actuals - mean(unlist(actuals)))^2) # Total Sum of Squares
-ss_residual <- sum((actuals - predictions)^2) # Residual Sum of Squares
-r_squared <- 1 - (ss_residual / ss_total)
-
-# Print the LOOCV R-squared
-cat("LOOCV R-squared:", r_squared, "\n")

### 2025-04-15 paper/paper.qmd
123

@@
 ---
-title: "Exploring Determinants of Child Mortality Globally (2022): The Role of Vaccination, Healthcare Spending, and Nutrition"
-subtitle: "Findings Reveal Vaccination and Healthcare Spending as Key Drivers of Mortality Reduction"
+title: "Improving Strong Gravitational Lensing Detection with Active Learning and Tailored Loss Functions Under Data Scarcity"
 author: 
   - Junbo Li
+  - Supervisor: Gwendolyn Eadie, Mike Walmsley
 thanks: "Code and data are available at: [https://github.com/Junbo345/Mortality_analysis](https://github.com/Junbo345/Mortality_analysis)."
@@
 date-format: long
-abstract: "Child mortality has significantly decreased over the past three decades, but progress remains uneven across regions.
-This study examines the under-five mortality rates of 2022 for each country, focusing on the effects of DPT vaccination coverage, health expenditure per capita, and food production indices.
-Using a Bayesian multi-linear modeling approach, we found that increased vaccination coverage and healthcare spending significantly reduce child mortality, while food production had no direct impact.
-These findings underscore the need for targeted interventions in high-mortality regions, suggesting the importance of vaccination programs and healthcare funding in achieving global child survival goals."
+abstract: "Strong gravitational lensing is a rare yet critical phenomenon in astrophysics, offering insights into dark matter, galaxy evolution, and cosmology.
+However, its scarcity in observational data presents a significant challenge for machine learning models, particularly in supervised learning frameworks that rely on large amounts of labeled examples.
+In this study, we propose an approach that integrates active learning with specialized loss functions to improve the detection of strong lensing events under limited data conditions.
+We employ the BADGE active learning strategy to select the most informative samples for labeling and evaluate two custom loss functions—Focal Loss and RankNet Loss—against a baseline model using standard cross-entropy.
+Our experiments reveal that while Focal Loss shows modest improvements in surfacing true lensing candidates, its computational cost may limit practical application, especially in active learning contexts.
+RankNet Loss, on the other hand, suffers from pairing limitations due to class imbalance.
+We also introduce a more task-aligned evaluation metric based on the number of correctly identified lensing cases in the top 10% of ranked predictions.
+While no single method decisively outperforms the others, our results offer useful guidance for balancing performance and efficiency in rare-event classification problems and lay the groundwork for future enhancements through better feature learning, metric design, and data acquisition strategies."
 format: pdf
@@
 To overcome this challenge, we adopt a state-of-the-art active learning strategy known as BADGE, which selects the most informative data points for human annotation.
-This method aims to improve training efficiency by prioritizing samples that are both uncertain and diverse [@cite].
-BADGE has shown strong performance in other low-label settings and is recognized as one of the most effective techniques in active learning [@cite].
+This method aims to improve training efficiency by prioritizing samples that are both uncertain and diverse [@ash2020deep].
+BADGE has shown strong performance in other low-label settings and is recognized as one of the most effective techniques in active learning [@ash2020deep].
 

### 2025-04-15 paper/paper.qmd
Ref added

@@
   - Supervisor: Gwendolyn Eadie, Mike Walmsley
-thanks: "Code and data are available at: [https://github.com/Junbo345/Mortality_analysis](https://github.com/Junbo345/Mortality_analysis)."
+thanks: "Code and data are available at: [https://github.com/Junbo345/Actinve_learning_Strong_Lensing](https://github.com/Junbo345/Actinve_learning_Strong_Lensing)."
 date: today

### 2025-04-15 paper/paper.qmd
Author

--- 2025-04-15 paper/paper.qmd
@@
 author: 
-  - Junbo Li
-  - Supervisor: Gwendolyn Eadie, Mike Walmsley
-thanks: "Code and data are available at: [https://github.com/Junbo345/Mortality_analysis](https://github.com/Junbo345/Mortality_analysis)."
+  - name: "Junbo Li"
+  - name: "Gwendolyn Eadie"
+    role: "Supervisor"
+  - name: "Mike Walmsley"
+    role: "Supervisor"
+thanks: "Code and data are available at: [https://github.com/Junbo345/Actinve_learning_Strong_Lensing](https://github.com/Junbo345/Actinve_learning_Strong_Lensing)."
 date: today
@@
 Section @sec-results presents the performance results under the proposed method.
-Finally, Section @sec-diss discusses the implications of our findings, current limitations, and directions for future research in detecting strong lensing events.
+Finally, Section @sec-discussion discusses the implications of our findings, current limitations, and directions for future research in detecting strong lensing events.
 
@@
 
-# Results {#sec-Results}
+# Results {#sec-results}
 

### 2025-04-16 paper/paper.qmd
reference added

@@
 
-Strong gravitational lensing occurs when a massive foreground object, such as a galaxy cluster, bends the light from a more distant source, producing distinctive arcs or rings in astronomical images.
-Although these phenomena are of great importance to cosmology, they are extremely rare—accounting for only a small fraction of all astronomical observations.
-Traditional supervised learning approaches typically require large amounts of labeled data, which is not feasible in this context due to the scarcity of strong lensing examples.
+Strong gravitational lensing occurs when a massive foreground object, such as a galaxy cluster, bends the light from a more distant source, producing distinctive arcs or rings in astronomical images [@2010_Treu].
+Although these phenomena are of great importance to cosmology, they are extremely rare—accounting for only a small fraction of all astronomical observations [@suyu2024strong].
+Traditional supervised learning approaches typically require large amounts of labeled data [@hsu2025leveraging], which is not feasible in this context due to the scarcity of strong lensing examples.
 
@@
 Features are first extracted using a convolutional neural network (CNN), and then concatenated with expert-provided labels for training.
-The resulting dataset contains approximately ~ total samples, with ~ representing strong lensing cases.
+The resulting dataset contains approximately 3329 total samples, with 255 representing strong lensing cases.
 
